#Load packages
```{r}
#install.packages("gbm")
library('gbm') #generalized boosted models
library('caTools') #calculates AUC
#install.packages("ROCR")
library('ROCR') #makes ROC curves
#install.packages("dismo")
library('dismo')
library('caret')
library("dataRetrieval")
library("pdp")          # model visualization
library("ggplot2")      # model visualization
library("lime")         # model visualization
#install.packages("pROC")
library("pROC")
#install.packages("e1071", repos="http://R-Forge.R-project.org")
library("e1071")
library( "MASS" )    #    used to generate correlated variables
library("sp")
library("dplyr")
library("Hmisc")    #   used for graphing se bars
#install.packages("rnoaa")
library("rnoaa")
```

#Read water quality data, remove -9999 and NA
```{r}
c<-"/Users/owner/Desktop/crypto/ec"
setwd(c)
cc<-read.csv("narrowresult-EC_PA_WI_OH.tsv",header=T,sep="\t")
dim(cc)
cc<-cc[,c("ActivityStartDate","MonitoringLocationIdentifier"
          #,"ResultDetectionConditionText"
          ,"CharacteristicName"
          #,"ResultSampleFractionText"
          , "ResultMeasureValue", "ResultMeasure.MeasureUnitCode"
          #,"ResultCommentText","ResultLaboratoryCommentText"
          )]
cc$ResultMeasureValue<-as.numeric(as.character(cc$ResultMeasureValue))
head(cc)
summary(cc$ResultMeasureValue)
cc<-cc[!(cc$ResultMeasureValue==-9999.0| is.na(cc$ResultMeasureValue)),]
summary(cc$ResultMeasureValue)
save(cc,file="/Users/owner/Desktop/crypto/ec/cc.Rdata")
```

#read stations
```{r}
c_s<-"/Users/owner/Desktop/crypto/ec"
setwd(c_s)
cc_s<-read.csv("station-EC_WI_PA_OH.csv",header=T,sep=',')
#cc_s2<-read.csv("station201909.csv",header=T,sep=',')
head(cc_s)
dim(cc_s)
cc_s<-cc_s[,c("MonitoringLocationIdentifier","DrainageAreaMeasure.MeasureValue","DrainageAreaMeasure.MeasureUnitCode", "ContributingDrainageAreaMeasure.MeasureValue","ContributingDrainageAreaMeasure.MeasureUnitCode", "LatitudeMeasure", "LongitudeMeasure")]
head(cc_s)
#cc_s2<-cc_s2[,c("MonitoringLocationIdentifier","DrainageAreaMeasure.MeasureValue","DrainageAreaMeasure.MeasureUnitCode", "ContributingDrainageAreaMeasure.MeasureValue","ContributingDrainageAreaMeasure.MeasureUnitCode", "LatitudeMeasure", "LongitudeMeasure")]
save(cc_s,file="/Users/owner/Desktop/crypto/ec/cc_s.Rdata")
```

#find USGS "river" stations
```{r}
USGS_cc<-cc[grepl("USGS", cc$MonitoringLocationIdentifier)==T,]   
dim(USGS_cc)
USGS_cc$USGS<-USGS_cc$MonitoringLocationIdentifier

USGS_cc$USGS<-gsub('USGS-','',as.character(USGS_cc$USGS))
USGS_cc3<-USGS_cc[nchar(USGS_cc$USGS)==8,]

USGS_cc2<-merge(USGS_cc3,cc_s,by.x="MonitoringLocationIdentifier",by.y="MonitoringLocationIdentifier")

dim(USGS_cc)
dim(USGS_cc2)
dim(USGS_cc3)
head(USGS_cc2)
save(USGS_cc2,file="/Users/owner/Desktop/crypto/ec/USGS_cc2.Rdata")
```

#get Q (LULC first? Q takes too long)
```{r}

USGS_cc2$Q<-" "
for (i in c(1:length(USGS_cc2$USGS)))

{
  
siteNo <- as.character(USGS_cc2[i,]$USGS)
pCode <- "00060" #discharge
#pCode shortName
#00060 Discharge [ft3/s]
#00065 Gage height [ft]
#00010 Temperature [C]
#00045 Precipitation [in]
#00400 pH
start.date <- as.character(USGS_cc2[i,]$ActivityStartDate)
end.date <- as.character(USGS_cc2[i,]$ActivityStartDate)
if (nrow(readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = start.date,
                     endDate = end.date))>0)
{
  USGS_cc2[i,]$Q<- readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = start.date,
                     endDate = end.date)$X_00060_00003
}
else
{
 USGS_cc2[i,]$Q<-NA 
}
  }
head(USGS_cc2)

```

#get runoff (negative Q aND RUNOFF!) (USGS-04194085)
```{r}
USGS_cc4<-USGS_cc2[is.na(USGS_cc2$Q)==F,]
USGS_cc4$runoff<-""
#ft s/sq mi
#1 mi = 5280 ft.
USGS_cc4$runoff<-(as.numeric(USGS_cc4$Q)*86400)/((USGS_cc4$DrainageAreaMeasure.MeasureValue)*5280*5280)*(0.3048*1000) #mm/d
summary(USGS_cc4$runoff)

USGS_cc4<-USGS_cc4[!USGS_cc4$MonitoringLocationIdentifier=="USGS-04194085",]

 missQ<-as.data.frame(table(USGS_cc2[!(USGS_cc2$MonitoringLocationIdentifier %in% USGS_cc4$MonitoringLocationIdentifier),]$MonitoringLocationIdentifier))

save(USGS_cc4,file="/Users/owner/Desktop/crypto/ec/USGS_cc4_20190507.Rdata")
```

#load the data
```{r}
USGS_cc4_20190507<-load("/Users/owner/Desktop/crypto/ec/USGS_cc4_20190507.Rdata")
head(USGS_cc4_20190507)
head(USGS_cc4)

```

#remove USGS-04194085
```{r}
dim(USGS_cc4)

USGS_cc4<-USGS_cc4[!USGS_cc4$MonitoringLocationIdentifier=="USGS-04194085",]

dim(USGS_cc4)

summary(USGS_cc4$runoff)

```

#load LULC data
```{r}
CntlPlains_cry_ws<-load("/Users/owner/Desktop/crypto/bas_ws/bas_nonref_CntlPlains_cry_ws/bas_nonref_CntlPlains_cry_ws.Rdata")
head(CntlPlains_cry_ws)
EastHghlnds_cry_ws<-load("/Users/owner/Desktop/crypto/bas_ws/bas_nonref_EastHghlnds_cry_ws/bas_nonref_EastHghlnds_cry_ws.Rdata")
head(EastHghlnds_cry_ws)
NorthEast_cry_ws<-load("/Users/owner/Desktop/crypto/bas_ws/bas_nonref_NorthEast_cry_ws/bas_nonref_NorthEast_cry_ws.Rdata")
head(NorthEast_cry_ws)
head(bas_nonref_NorthEast_cry_ws)
```

#merge the lulc
```{r}
USGS_cc4_lulc_CntlPlains<-merge(bas_nonref_CntlPlains_cry_ws,USGS_cc4,by.x="GAGE_ID", by.y="MonitoringLocationIdentifier")
USGS_cc4_lulc_EastHghlnds<-merge(bas_nonref_EastHghlnds_cry_ws,USGS_cc4,by.x="GAGE_ID", by.y="MonitoringLocationIdentifier" )
USGS_cc4_lulc_NorthEast<-merge(bas_nonref_NorthEast_cry_ws,USGS_cc4,by.x="GAGE_ID", by.y="MonitoringLocationIdentifier" )

USGS_cc4_lulc_CntlPlains_EastHghlnds_cry_ws<-rbind(USGS_cc4_lulc_CntlPlains,USGS_cc4_lulc_EastHghlnds)
USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws<-rbind(USGS_cc4_lulc_NorthEast,USGS_cc4_lulc_CntlPlains_EastHghlnds_cry_ws)

dim(USGS_cc4_lulc_CntlPlains)
dim(USGS_cc4_lulc_CntlPlains_EastHghlnds_cry_ws)
dim(USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws)

save(USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws,file="/Users/owner/Desktop/crypto/ec/USGS_cc4_lulc_all_ws.Rdata")

```

```{r}
USGS_cc4_lulc_all_ws<-load("/Users/owner/Desktop/crypto/ec/USGS_cc4_lulc_all_ws.Rdata")
#USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws
```

#find the nearest climate station for each USGS station
```{r}
USGS_cc4_lulc<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws
m<-matrix(c(USGS_cc4_lulc$LongitudeMeasure,USGS_cc4_lulc$LatitudeMeasure),ncol=2)
mu<-unique(m)

#To get a full and current list of stations
stations <- ghcnd_stations()
save(stations,file="/Users/owner/Desktop/crypto/ec/stations.Rdata")

stations<-stations[!(stations$id %in% c("USC00103760","USC00331905","USC00361920","USC00365902","USC00366233","USC00368668","USC00475474","USC00476764","USC00476922","USC00479046","USW00014813","USC00336949","USC00360861","USC00361726")),]
#the last year later than 2009
station_last<-stations[stations$last_year>2009 & stations$first_year<1989 ,]
#temperature
station_last_TMAX_TMIN<-station_last[station_last$element %in% c("TMAX"),]
sm<-matrix(c(station_last_TMAX_TMIN$longitude,station_last_TMAX_TMIN$latitude ),ncol=2)
smu<-unique(sm)
#USGS_cc4_lulc_all_ws
#sta<-c("")
lon<-as.numeric("")
lat<-as.numeric("")
USGS_lon<-as.numeric("")
df<-data.frame(
  lon,lat,USGS_lon)

station_last_TMAX_TMIN_nearest[station_last_TMAX_TMIN_nearest$USGS_lon==-88.0436111,]

for (i in c(1:dim(mu)[1]))
{
 # print(mu[i,])
  km <- spDistsN1(smu, #climate station
                  mu[i,], longlat=TRUE)
  i2<-which(km==min(km))
  sdf<-data.frame(smu[i2,1],smu[i2,2],mu[i,][1])
  colnames(sdf)<-c("lon", "lat","USGS_lon")
  df<-rbind(df,sdf)
}
summary(df) #list of weather stations (can be repeated)
```

#the table of NOAA & USGS
```{r}
dim(station_last_TMAX_TMIN)
station_last_TMAX_TMIN_nearest<-merge(station_last_TMAX_TMIN, #NOAA stations
                                      df[-1,], #USGS lon  and the nearest NOAA stations
                                      by.x="latitude",by.y="lat")
station_last_TMAX_TMIN_nearest<-station_last_TMAX_TMIN_nearest[!station_last_TMAX_TMIN_nearest$state=="OR",]
# df2<-df
# colnames(df2)<-c("latitude","longitude","USGS_lon")
# df2<-df2[-1,]
# station_last_TMAX_TMIN_nearest<- full_join(station_last_TMAX_TMIN, #NOAA stations
#                                             df2, #USGS lon  and the nearest NOAA stations
#                                            by = c("latitude","longitude"))

#station_last_TMAX_TMIN_nearest <- station_last_TMAX_TMIN_nearest[!is.na(station_last_TMAX_TMIN_nearest$USGS_lon),]

dim(station_last_TMAX_TMIN_nearest)
table(station_last_TMAX_TMIN_nearest$element)

tmax_sta<-station_last_TMAX_TMIN_nearest[station_last_TMAX_TMIN_nearest$element=="TMAX",]
tmin_sta<-station_last_TMAX_TMIN_nearest[station_last_TMAX_TMIN_nearest$element=="TMIN",]
tmax_sta[!(tmax_sta$id %in% tmin_sta$id),]

#station_last_TMAX_TMIN_nearest<-station_last_TMAX_TMIN_nearest[station_last_TMAX_TMIN_nearest$element=="TMIN",]
save(station_last_TMAX_TMIN_nearest,file="/Users/owner/Desktop/crypto/ec/station_last_TMAX_TMIN_nearest.Rdata")
```

#1. (USGS_cc4_lulc) LULC-date-ResultMeasureValue-drainage area-"LatitudeMeasure"- "LongitudeMeasure"-"Q"-"runoff"
#2. (station_last_TMAX_TMIN_nearest) "latitude"-"id"-"longitude"-"elevation"-"state"-"name"-"gsn_flag"-"wmo_id"-"element"- "first_year" "last_year"-"lon"-"USGS_lon"
#3. merge by "USGS_lon"
```{r}
dim(USGS_cc4_lulc)
USGS_cc4_lulc_nearest<-merge(USGS_cc4_lulc,station_last_TMAX_TMIN_nearest,by.x= "LongitudeMeasure" ,by.y=  "USGS_lon")
dim(USGS_cc4_lulc[!(USGS_cc4_lulc$LongitudeMeasure %in% station_last_TMAX_TMIN_nearest$USGS_lon ),])
table(USGS_cc4_lulc[!(USGS_cc4_lulc$LongitudeMeasure %in% station_last_TMAX_TMIN_nearest$USGS_lon ),]$USGS)
dim(USGS_cc4_lulc_nearest)
save(USGS_cc4_lulc_nearest,file="/Users/owner/Desktop/crypto/ec/USGS_cc4_lulc_nearest.Rdata")
# id - ResultMeasureValue - RUNOFF- DATE - DrainageAreaMeasure.MeasureValu- LULC
```

#get tmin 
```{r}
u_sta<-unique( USGS_cc4_lulc_nearest$id)

# u_sta_d<-as.data.frame(u_sta)
# # 
# u_sta_d<-u_sta_d[!(u_sta_d$u_sta%in%c("USC00103760","USC00331905","USC00361920","USC00365902","USC00366233","USC00368668","USC00475474","USC00476764","USC00476922","USC00479046","USW00014813")),]
# # 
# u_sta<-as.character(u_sta_d)

id<-c("")
 tmin<-as.numeric("")
 date<-as.Date("1234-12-21")     
 mflag<-c("") 
 qflag<-c("") 
 sflag<-c("")
 sta_df<-data.frame(id,tmin,date,mflag,qflag,sflag)
 sta_df$id<-as.character(sta_df$id)
 
 for (s in c(1:length(u_sta)))
{
  u_sta[s]
  g<-ghcnd_search(u_sta[s], var = "TMIN")
  sta_tmin<-g$tmin
   sta_df<-rbind( sta_df,  sta_tmin)
 }

# length(sta_df$ActivityStartDate)
# sta_df<-sta_df[   sta_df$ActivityStartDate>as.Date("1989-01-01") & sta_df$ActivityStartDate<as.Date("2009-12-31"),]
# length(sta_df$ActivityStartDate)
#  
# table(sta_df[is.na(sta_df$tmin),]$id)

#USC00103760 USC00331905 USC00361920 USC00365902 USC00366233 USC00368668 
#        333           5           1          10          41           3 
#USC00475474 USC00476764 USC00476922 USC00479046 USW00014813 
#         21          53          81           3          93 

colnames(sta_df)<-c("id","tmin" ,"ActivityStartDate","mflag","qflag","sflag")    
USGS_cc4_lulc_nearest$ActivityStartDate<-as.Date(USGS_cc4_lulc_nearest$ActivityStartDate) 
sta_df$id<-as.character(sta_df$id)
USGS_cc4_lulc_nearest_tmin<- full_join(USGS_cc4_lulc_nearest, sta_df,by = c("id", "ActivityStartDate"))
USGS_cc4_lulc_nearest_tmin_rmna<-USGS_cc4_lulc_nearest_tmin[!is.na(USGS_cc4_lulc_nearest_tmin$ResultMeasureValue),]
dim(USGS_cc4_lulc_nearest_tmin_rmna)
USGS_cc4_lulc_nearest_tmin_rmna<-USGS_cc4_lulc_nearest_tmin_rmna[USGS_cc4_lulc_nearest_tmin_rmna$element=="TMAX",]
dim(USGS_cc4_lulc_nearest_tmin_rmna)
save(USGS_cc4_lulc_nearest_tmin_rmna,file="/Users/owner/Desktop/crypto/ec/USGS_cc4_lulc_nearest_tmin.Rdata")

```

#get tmax 
```{r}
id<-c("")
 tmax<-as.numeric("")
 date<-as.Date("1234-12-21")     
 mflag<-c("") 
 qflag<-c("") 
 sflag<-c("")
 sta_df_max<-data.frame(id,tmax,date,mflag,qflag,sflag)
 
 for (s in c(1:length(u_sta)))
{
  u_sta[s]
  g<-ghcnd_search(u_sta[s], var = "TMAX")
  sta_tmax<-g$tmax
   sta_df_max<-rbind( sta_df_max,  sta_tmax)
 }

colnames(sta_df_max)<-c("id","tmax" ,"ActivityStartDate","mflag","qflag","sflag")    
sta_df_max$id<-as.character(sta_df_max$id)


USGS_cc4_lulc_nearest_tmin_rmna_tmax<- full_join(USGS_cc4_lulc_nearest_tmin_rmna, sta_df_max,by = c("id", "ActivityStartDate"))
summary(USGS_cc4_lulc_nearest_tmin_rmna_tmax)

USGS_cc4_lulc_nearest_tmin_rmna_tmax<-USGS_cc4_lulc_nearest_tmin_rmna_tmax[!is.na(USGS_cc4_lulc_nearest_tmin_rmna_tmax$ResultMeasureValue),]

save(USGS_cc4_lulc_nearest_tmin_rmna_tmax,file="/Users/owner/Desktop/crypto/ec/USGS_cc4_lulc_nearest_tmin_rmna_tmax.Rdata")

```

#hysep function
```{r}

hysep <- function(Flow, Dates, Start=NULL, End=NULL, da,
                 select="sliding", STAID="Unknown") {
  ## Start of code: initial processing
  STAID <- as.character(STAID[1L])
  if(is.null(Start))
    Start <- Dates[1L]
  else if(is.character(Start))
    Start <- as.Date(Start)
  if(is.null(End))
    End <- Dates[length(Dates)]
  else if(is.character(End))
    End <- as.Date(End)
  sel <- (Dates >= Start) & (Dates <= End)
  Dates <- Dates[sel]
  Flow <- pmax(Flow[sel], 1e-99) # Convert 0 to a small number
  if(any(is.na(Flow)))
    stop("Missing values between ", Start, " and ", End)
  if(any(diff(as.double(Dates)) != 1))
    stop("Date data are not continuous between Start and End")
  select <- match.arg(select, c("sliding", "local minimum", "fixed"))
  Nact <- max(da^0.2, 1)
  N2star <- max((((2*Nact) ) %/% 2)*2 + 1, 3)
  ## Set up for fixed--construct intervals of length N2star
  Nobs <- length(Flow)
  Ngrp <- ceiling(Nobs / N2star)
  Grps <- inverse.rle(list(lengths=rep(N2star, Ngrp), values=seq(Ngrp)))
  length(Grps) <- Nobs # Truncate if necessary
  ## Compute the fixed method
  Mins <- tapply(Flow, Grps, min)
  Fixed <- Mins[as.character(Grps)]
  ## Now the sliding method
  Slide <- sapply(seq(N2star, Nobs), function(i)
    min(Flow[seq(i - N2star + 1L, i)])
  )
  SlB <- Slide[1L]
  SlE <- Slide[length(Slide)]
  Nfil <- (N2star - 1L) / 2L
  Slide <- c(rep(SlB, Nfil), Slide, rep(SlE, Nfil))
  ## And the local minimum
  Mid <- as.integer((N2star) / 2)
  LocMin <- sapply(seq(N2star, Nobs), function(i)
    min(Flow[seq(i - N2star + 1L, i)]) == Flow[i - Mid]
  )
  LocMin <- c(rep(FALSE, Nfil), LocMin, rep(FALSE, Nfil))
  ## Need to trap short periods where only 1 local minimum
  if(sum(LocMin) == 1L) {
    warning("Only one local minimum in calibration period")
    LocMin <- pmax(Flow[LocMin], 0.01)
  } else
    LocMin <- exp(approx(which(LocMin), log(pmax(Flow[LocMin], 0.01)), xout=seq(Nobs), rule=2)$y)
  LocMin <- pmin(Flow, LocMin) # recover 0s and tails
  if(select == "fixed")
    BaseQ <- Fixed
  else if(select == "sliding")
    BaseQ <- Slide
  else 
    BaseQ <- LocMin
  retval <- data.frame(Dates=Dates, BaseQ=round(BaseQ, 3L), 
                       Flow=Flow, Fixed=Fixed, Sliding=Slide, 
                       LocalMin=LocMin)
  if(!is.null(STAID))
    attr(retval, "STAID") <- STAID
  attr(retval, "type") <- "hysep"
  class(retval) <- c("baseflow", "data.frame")
  return(retval)
}

```

#need to get Q in previous 44 days 
```{r}
USGS_cc4<-USGS_cc4_lulc_nearest_tmin_rmna_tmax
USGS_cc4$storm<-" "
for (i in c(1:length(USGS_cc4$USGS)))
#for (i in c(41:59))
{
siteNo <- as.character(USGS_cc4[i,]$USGS)
pCode <- "00060" #discharge
start.date <- as.character(as.Date(USGS_cc4[i,]$ActivityStartDate,format=("%Y-%m-%d"))-44)
end.date <- as.character(USGS_cc4[i,]$ActivityStartDate)

if (nrow(readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = start.date,
                     endDate = end.date))>0)
{
  print(nrow(readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = start.date,
                     endDate = end.date)))
  USGS_cc4[i,]$Q<- readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = end.date,
                     endDate = end.date)$X_00060_00003
  
 st<- paste("Q_44days",i,sep='_')
  st<- readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = start.date,
                     endDate = end.date)
  print (paste("Q_44days",i,sep='_'))
  #hysep(s1372500$X105182_00060_00003,Dates=2/1/2019,da=1)
  #head(s1372500)
  #agency_cd site_no datetime X105182_00060_00003 X105182_00060_00003_cd
#1      USGS 1372500 2/1/2019                 399                      P
test<-hysep(as.integer(st$X_00060_00003),Dates=
              #2008-09-27
              #9/27/2008
              st$Date
            ,da=1)
print(test)
USGS_cc4[i,]$storm<-(1-test$BaseQ[45]/test$Flow[45])
}
else
{
 USGS_cc4[i,]$storm<-NA 
 USGS_cc4[i,]$Q<-NA 
}
print(paste("day ",i," done",sep=','))
  }
head(USGS_cc4)
plot(USGS_cc4$storm,as.numeric(USGS_cc4$runoff))
                                
save(USGS_cc4,file="/Users/owner/Desktop/crypto/ec/USGS_cc4_20190507_2.Rdata")
```

```{r}
load("/Users/owner/Desktop/crypto/ec/USGS_cc4_20190507_2.Rdata")
#usgs_cc4 (USGS_lon_date)
USGS_cc4$storm<-as.numeric(USGS_cc4$storm)
summary(USGS_cc4$storm)
USGS_cc4<-USGS_cc4[!is.na(USGS_cc4$storm),]
USGS_cc4<-USGS_cc4[!is.na(USGS_cc4$tmax),]
USGS_cc4$Q<-as.numeric(USGS_cc4$Q)
summary(USGS_cc4)
#"storm", "tmax"
```


<!-- #get temperature from the nearest station -->
<!-- #https://cran.r-project.org/web/packages/rnoaa/rnoaa.pdf -->
<!-- ```{r} -->
<!-- head(USGS_cc4) -->
<!-- #TEST -->
<!-- #coops_search(station_name = 8723970, begin_date = 20120301,end_date = 20141001, datum = "stnd", product = "monthly_mean") -->
<!-- #GET THE LIST OF STATION -->
<!-- #station_data <- ghcnd_stations()  -->
<!-- lat_lon_df<-USGS_cc4[, c("MonitoringLocationIdentifier", "LatitudeMeasure",  "LongitudeMeasure")  ] -->
<!-- colnames(lat_lon_df)<-c("id","latitude", "longitude") -->
<!-- #save(station_data,file="/Users/owner/Desktop/CRYPTO_CODE/station_data.Rdata") -->
<!-- load("/Users/owner/Desktop/CRYPTO_CODE/station_data.Rdata") -->
<!-- nearby_stations <-  meteo_nearby_stations(lat_lon_df = lat_lon_df, -->
<!--                     station_data = station_data, radius = 10, var = c("PRCP", "TMAX","TAVG","TMIN")) -->
<!-- head(nearby_stations) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- temperature<-read.csv("/Users/owner/Desktop/crypto/cry_station/1725692.csv",header=T,sep=',') -->
<!-- temperature -->
<!-- temperature<-temperature[temperature$STATION=="USW00014895",] -->

<!-- #TAVG<-temperature[,c("DATE","TAVG")] -->
<!-- TAVG<-temperature[,c("DATE","TAVG","TMAX","TMIN")] -->
<!-- w<-which(is.na(TAVG$TAVG)==T) -->
<!-- TAVG[w,]$TAVG<-(TAVG[w,]$TMIN+TAVG[w,]$TMAX)/2 -->
<!-- summary(TAVG$TAVG) -->
<!-- #load("/Users/owner/Desktop/crypto/USGS_cc4.Rdata") -->

<!-- USGS_cc4_temp<-merge(USGS_cc4,TAVG,by.x="ActivityStartDate", by.y="DATE" ) -->
<!-- USGS_cc4<-USGS_cc4_temp -->
<!-- save(USGS_cc4,file="/Users/owner/Desktop/crypto/USGS_cc4_20190507_2.Rdata") -->

<!-- ``` -->

<!-- ```{r} -->
<!-- USGS_cc4_20190507_2<-load("/Users/owner/Desktop/crypto/USGS_cc4_20190507_2.Rdata") -->
<!-- head(USGS_cc4_20190507_2) -->
<!-- colnames(USGS_cc4) -->
<!-- ``` -->




<!-- ```{r} -->
<!-- no_LULC<-(USGS_cc4[!(USGS_cc4$MonitoringLocationIdentifier %in% bas_nonref_CntlPlains_cry_ws$GAGE_ID) & !(USGS_cc4$MonitoringLocationIdentifier %in% bas_nonref_EastHghlnds_cry_ws$GAGE_ID) & !(USGS_cc4$MonitoringLocationIdentifier %in% bas_nonref_NorthEast_cry_ws$GAGE_ID) ,]) -->

<!-- no_LULC_t<-as.data.frame(table(no_LULC$MonitoringLocationIdentifier)) -->

<!-- ``` -->

<!-- #USGS-01562000 -->
<!-- ```{r} -->
<!-- #GAGE_ID_01562000.shp is in bas_nonref_EastHghlnds_cry_ws -->
<!-- ``` -->


<!-- ```{r} -->
<!-- load("/Users/owner/Desktop/crypto/bas_ws/USGS_cc4_lulc_all_ws.Rdata") -->
<!-- ``` -->


<!-- #lump lulc -->
<!-- ```{r} -->
<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$Water<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc11 -->

<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$Developed<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc21+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc22+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc23+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc24 -->

<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$Barren<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc31 -->

<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$Forest<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc41+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc42+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc43 -->

<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$Shrubland<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc52 -->

<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$Herbaceous<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc71 -->

<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$Cultivated<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc81+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc82 -->
<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$Wetland<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc90+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc95 -->

<!-- head(USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws) -->

<!-- save(USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws,file="/Users/owner/Desktop/crypto/bas_ws/USGS_cc4_lulc_lumped_all_ws.Rdata") -->

<!-- ``` -->


<!-- #lump lulc (put lulc with mean<10% to "other") -->
<!-- ```{r} -->
<!-- #other -->
<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$Other<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc11+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc31+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc52+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc71+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc90+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc95 #water #Barren # Shrubland   #Herbaceous #wetland -->

<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$Developed<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc21+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc22+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc23+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc24 -->

<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$Forest<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc41+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc42+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc43 -->

<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$Cultivated<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc81+USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$lulc82 -->

<!-- head(USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws) -->

<!-- #make sure the sum is 100 -->
<!-- sum(USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws[1,c("Other","Developed" -->
<!-- #,"Barren" -->
<!-- ,"Forest", -->
<!--          #                                                    "Shrubland", -->
<!--                                                              #"Herbaceous", -->
<!--                                                              "Cultivated" -->
<!-- #                                                           ,  "Wetland" -->
<!-- )]) -->

<!-- save(USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws,file="/Users/owner/Desktop/crypto/bas_ws/USGS_cc4_lulc_lumped_all_ws.Rdata") -->

<!-- ``` -->

<!-- ```{r} -->
<!-- load("/Users/owner/Desktop/crypto/bas_ws/USGS_cc4_lulc_lumped_all_ws.Rdata") -->
<!-- head(USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws) -->
<!-- ``` -->

<!-- #rename columns -->
<!-- ```{r} -->
<!-- USGS_cc4_lulc<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws -->

<!-- colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="TAVG"] <- "Temperature" -->
<!-- colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="storm"] <- "Surface_runoff" -->
<!-- colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="detect"] <- "Detect" -->
<!-- colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="DrainageAreaMeasure.MeasureValue"] <- "Watershed_Area" -->
<!-- colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="runoff"] <- "Runoff" -->
<!-- colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="Other"] <- "Other_LULC" -->
<!-- ``` -->

<!-- #add interactive terms -->
<!-- ```{r} -->
<!-- USGS_cc4_lulc_i<-USGS_cc4_lulc -->
<!-- USGS_cc4_lulc_i$Developed_Surface_runoff<-USGS_cc4_lulc_i$Developed*USGS_cc4_lulc_i$Surface_runoff -->
<!-- USGS_cc4_lulc_i$Developed_Runoff<-USGS_cc4_lulc_i$Developed*USGS_cc4_lulc_i$Runoff -->
<!-- USGS_cc4_lulc_i$Forest_Surface_runoff<-USGS_cc4_lulc_i$Forest*USGS_cc4_lulc_i$Surface_runoff -->
<!-- USGS_cc4_lulc_i$Forest_Runoff<-USGS_cc4_lulc_i$Forest*USGS_cc4_lulc_i$Runoff -->
<!-- USGS_cc4_lulc_i$Runoff_Surface_runoff<-USGS_cc4_lulc_i$Runoff*USGS_cc4_lulc_i$Surface_runoff -->
<!-- USGS_cc4_lulc_i$Temperature_Surface_runoff<-USGS_cc4_lulc_i$Temperature*USGS_cc4_lulc_i$Surface_runoff -->
<!-- ``` -->

#binary detect
```{r}
USGS_cc4_lulc<-USGS_cc4
dim(USGS_cc4_lulc)

USGS_cc4_lulc$Detect<-as.numeric(0)
USGS_cc4_lulc[USGS_cc4$ResultMeasureValue>235,]$Detect<-as.numeric(1)
dim(USGS_cc4_lulc)

USGS_cc4_lulc$Developed<-USGS_cc4_lulc$lulc21    + USGS_cc4_lulc$lulc22+USGS_cc4_lulc$lulc23+USGS_cc4_lulc$lulc24

USGS_cc4_lulc$Forest <-USGS_cc4_lulc$lulc41   + USGS_cc4_lulc$lulc42 +USGS_cc4_lulc$lulc43

USGS_cc4_lulc<-USGS_cc4_lulc[!is.na(USGS_cc4_lulc$tmin),]
USGS_cc4_lulc$tmin<-USGS_cc4_lulc$tmin/10
USGS_cc4_lulc$tmax<-USGS_cc4_lulc$tmax/10
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="storm"] <- "Surface_runoff"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="tmin"] <- "Min_temperature"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="tmax"] <- "Max_temperature"
```

#split the data
```{r}

#Partitioning the data and running a GBM

#Creating training and test sets
set.seed(123) #uses a fixed seed so that every call will result in the same output.

USGS_cc4_lulc<-USGS_cc4_lulc[,c(52:53,15:17,12,7:9,3   #LULC
                                ,42,46 #TEMPERATURE
                                ,50 #SURFACE Q
                                ,29 #RUNOFF
                                ,23 #WATERSHED AREA
                                ,51 #DETECT
                                )]

intrain<-createDataPartition(y=USGS_cc4_lulc$Detect ,
                             p=0.8,
                             list=FALSE)
#class(intrain) "matrix"

test<-USGS_cc4_lulc[-intrain,]
training<-USGS_cc4_lulc[intrain,]

dim(test)
dim(training)
```


<!-- #split the data (interactive) -->
<!-- ```{r} -->

<!-- set.seed(123) #uses a fixed seed so that every call will result in the same output. -->

<!-- USGS_cc4_lulc_i<-USGS_cc4_lulc_i[,c(41:44 #LULC -->
<!--                                 ,38 #TEMPERATURE -->
<!--                                 ,37 #SURFACE Q -->
<!--                                 ,36 #RUNOFF -->
<!--                                 ,29 #WATERSHED AREA -->
<!--                                 ,26 #DETECT -->
<!--                                 ,45:50 #interactive -->
<!--                                 )] -->

<!-- intrain<-createDataPartition(y=USGS_cc4_lulc_i$Detect , -->
<!--                              p=0.8, -->
<!--                              list=FALSE) -->
<!-- #class(intrain) "matrix" -->

<!-- test<-USGS_cc4_lulc_i[-intrain,] -->
<!-- training<-USGS_cc4_lulc_i[intrain,] -->

<!-- dim(test) -->
<!-- dim(training) -->
<!-- ``` -->

#the model
```{r}
model<-as.formula(paste(colnames(USGS_cc4_lulc)[16], "~",      paste(colnames(USGS_cc4_lulc)[c(1:15)], collapse="+"), #families
                        sep = ""))
#that error message typically means you don’t have values for a variable represented in the model building process. 
water_gbm_lumped<- gbm(model,
             data=training, 
             distribution="bernoulli",
             n.trees=400,
             shrinkage=0.1,
             interaction.depth=1,
             bag.fraction=0.50,
             #train.fraction=1,
             n.minobsinnode=10,
            cv.folds=10, # that error occurs when you have variable levels that exist in the test set which don't exist in the training set.
             keep.data=TRUE,
             class.stratify.cv=TRUE,
             verbose="CV")
#jpeg(paste("/Users/owner/Desktop/crypto/ec/","lumped_perf",'.jpg',sep=''))
gbm.perf(water_gbm_lumped)

```

<!-- #the model (w interactive terms) -->
<!-- ```{r} -->
<!-- model<-as.formula(paste(colnames(USGS_cc4_lulc_i)[9], "~",      paste(colnames(USGS_cc4_lulc_i)[c(1:8,10:15)], collapse="+"), #families -->
<!--                         sep = "")) -->
<!-- #that error message typically means you don’t have values for a variable represented in the model building process.  -->
<!-- water_gbm_lumped_i<- gbm(model, -->
<!--              data=training,  -->
<!--              distribution="bernoulli", -->
<!--              n.trees=400, -->
<!--              shrinkage=0.1, -->
<!--              interaction.depth=2, -->
<!--              bag.fraction=0.50, -->
<!--              #train.fraction=1, -->
<!--              n.minobsinnode=10, -->
<!--              cv.folds=10, # that error occurs when you have variable levels that exist in the test set which don't exist in the training set. -->
<!--              keep.data=TRUE, -->
<!--              class.stratify.cv=TRUE, -->
<!--              verbose="CV") -->
<!-- #jpeg(paste("/Users/owner/Desktop/crypto/ec/","lumped_perf",'.jpg',sep='')) -->
<!-- gbm.perf(water_gbm_lumped_i) -->
<!-- p<-(water_gbm_lumped_i) -->
<!-- ``` -->



```{r}
best.iter<-gbm.perf(water_gbm_lumped)
gbm:::interact.gbm(water_gbm_lumped, data=USGS_cc4_lulc, i.var=c("Surface_runoff","Runoff"), best.iter)
gbm:::interact.gbm(water_gbm_lumped, data=USGS_cc4_lulc, i.var=c("Cultivated","Runoff"), best.iter)
```


#summary
```{r}
par(mar = c(5, 14, 1, 1))
# summary(
#  water_gbm_lumped , 
#   cBars = 10,
#   method = relative.influence, # also can use permutation.test.gbm
#   las = 2
#   )

summary(
 water_gbm_lumped , 
  cBars = 10,
  method = relative.influence, # also can use permutation.test.gbm
  las = 2
  )

#boxplot(USGS_cc4_lulc$lulc90~USGS_cc4_lulc$detect )
#boxplot(log(as.numeric(USGS_cc4_lulc$Surface_runoff )) ~USGS_cc4_lulc$Detect )

#percent positive in summer & winter
#length(USGS_cc4_lulc[USGS_cc4_lulc$detect==1&USGS_cc4_lulc$summer==1,1])/length(USGS_cc4_lulc[USGS_cc4_lulc$summer==1,1])
#length(USGS_cc4_lulc[USGS_cc4_lulc$detect==1&USGS_cc4_lulc$winter==1,1])/length(USGS_cc4_lulc[USGS_cc4_lulc$winter==1,1])
#print(water_gbm)

```

#n a ROC curve the true positive rate (Sensitivity) is plotted in function of the false positive rate (100-Specificity) for different cut-off points of a parameter.
```{r}
preds <- predict.gbm( water_gbm_lumped, test,n.trees=
                        #water_gbm$gbm.call$best.trees
                        class(gbm.perf(water_gbm_lumped)),
                      type="response")
calc.deviance(obs=as.numeric(test$Detect ) , pred=preds
              , calc.mean=TRUE
              )
roc_obj <- roc(test$Detect ,preds) #roc(response,predictor)
auc(roc_obj)
#jpeg(paste("/Users/owner/Desktop/crypto/ec/","roc_lumped",'.jpg',sep=''))
plot(roc_obj)
#dev.off()
plot(test$Detect ,preds)

predict_binary<-as.factor(ifelse(preds>0.5,1,0) )
test_obs<-as.factor(test$Detect)
confusionMatrix(predict_binary,test_obs)
```


#n a ROC curve the true positive rate (Sensitivity) is plotted in function of the false positive rate (100-Specificity) for different cut-off points of a parameter.
```{r}
preds <- predict.gbm( water_gbm_lumped, test,n.trees=
                        #water_gbm$gbm.call$best.trees
                        class(gbm.perf(water_gbm_lumped)),
                      type="response")
calc.deviance(obs=as.numeric(test$Detect ) , pred=preds
              , calc.mean=TRUE
              )
roc_obj <- roc(test$Detect ,preds) #roc(response,predictor)
auc(roc_obj)
#jpeg(paste("/Users/owner/Desktop/crypto/ec/","roc_lumped",'.jpg',sep=''))
plot(roc_obj)
#dev.off()
plot(test$Detect ,preds)

predict_binary<-as.factor(ifelse(preds>0.5,1,0) )
test_obs<-as.factor(test$Detect)
confusionMatrix(predict_binary,test_obs)
```

#Partial dependence plots. After the most relevant variables have been identified, the next step is to attempt to understand how the response variable changes based on these variables. 
#PDPs plot the change in the average predicted value as specified feature(s) vary over their marginal distribution. 
```{r}
jpeg(paste("/Users/owner/Desktop/crypto/ec/","relative_influence",'.jpg',sep=''))
par(mar = c(5, 11, 1, 1))
summary(
 water_gbm_lumped , 
  cBars = 10,
  method = relative.influence, # also can use permutation.test.gbm
  las = 2
  )
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/ec/","runoff",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "runoff", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
#+scale_y_continuous(labels = scales::dollar)
 #+scale_y_continuous(labels = scales::marginal effect )
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/ec/","Surface_runoff",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Surface_runoff", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/ec/","Max_temperature",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Max_temperature", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/ec/","Min_temperature",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Min_temperature", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/ec/","Runoff",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Runoff", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()



jpeg(paste("/Users/owner/Desktop/crypto/ec/","Developed",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Developed", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/ec/","Forest",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Forest", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/ec/","Cultivated",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Cultivated", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/ec/","Watershed_Area",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Watershed_Area", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
#partialPlot(boston.rf, pred.data = boston, x.var = "lstat",cex.lab=2.3)



```

#PDP for interactive terms
```{r}
jpeg(paste("/Users/owner/Desktop/crypto/ec/","relative_influence_i",'.jpg',sep=''))
par(mar = c(5, 14, 1, 1))
summary(
 water_gbm_lumped , 
  cBars = 10,
  method = relative.influence, # also can use permutation.test.gbm
  las = 2
  )
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Surface_runoff_i",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Surface_runoff", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Runoff_Surface_runoff_i",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Runoff_Surface_runoff", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Temperature_Surface_runoff_i",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Temperature_Surface_runoff", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Developed_Surface_runoff",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Developed_Surface_runoff", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Developed_Runoff",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Developed_Runoff", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Temperature_i",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Temperature", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Forest_Surface_runoff",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Forest_Surface_runoff", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Forest_Runoff",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Forest_Runoff", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Runoff_i",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Runoff", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Forest_i",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Forest", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Developed_i",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Developed", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Cultivated_i",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Cultivated", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Watershed_Area_i",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Watershed_Area", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
#partialPlot(boston.rf, pred.data = boston, x.var = "lstat",cex.lab=2.3)



```


```{r}
pretty.gbm.tree(water_gbm_lumped,i.tree=40)
```

```{r}
library(dismo)
data(Anguilla_train)
head(Anguilla_train)
angaus.tc5.lr01 <- gbm.step(data=Anguilla_train, gbm.x = 3:13, gbm.y = 2,
                        family = "bernoulli", tree.complexity = 5,                         learning.rate = 0.01, bag.fraction = 0.5)

angaus.tc5.lr005 <- gbm.step(data=Anguilla_train, gbm.x = 3:13, gbm.y = 2, 
                      family = "bernoulli", tree.complexity = 5,
                      learning.rate = 0.005, bag.fraction = 0.5)

find.int <- gbm.interactions(angaus.tc5.lr005)
find.int$interactions
```









#log
```{r}

USGS_cc4_lulc<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="TAVG"] <- "Temperature"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="storm"] <- "Surface_runoff"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc21"] <- "Developed_OpenSpace"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc22"] <- "Developed_LowIntensity"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc41"] <- "Forest"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc81"] <- "Pasture"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc82"] <- "Cultivated_Crops"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="detect"] <- "Detect"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="DrainageAreaMeasure.MeasureValue"] <- "Watershed_Area"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="runoff"] <- "log_runoff"

USGS_cc4_lulc_log<-USGS_cc4_lulc

USGS_cc4_lulc_log$log_runoff<-log(USGS_cc4_lulc_log$log_runoff)
#Partitioning the data and running a GBM

#Creating training and test sets
set.seed(123) #uses a fixed seed so that every call will result in the same output.

USGS_cc4_lulc_log<-USGS_cc4_lulc_log[,c(36:38,29,26,3:7)]

intrain<-createDataPartition(y=USGS_cc4_lulc_log$Detect ,
                             p=0.8,
                             list=FALSE)
#class(intrain) "matrix"

test<-USGS_cc4_lulc_log[-intrain,]
training<-USGS_cc4_lulc_log[intrain,]

dim(test)
dim(training)
```


<!-- #add summer and winter -->
<!-- ```{r} -->

<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$summer<-0 -->
<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws[USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$month==6|USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$month==7|USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$month ==8 ,]$summer<-1 -->
<!-- summary(USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$summer) -->

<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$winter<-0 -->
<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws[USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$month==12|USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$month==1|USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$month ==2 ,]$winter<-1 -->
<!-- summary(USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$winter) -->

<!-- ``` -->

#run the model
```{r}

USGS_cc4_lulc<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws

#Partitioning the data and running a GBM

#Creating training and test sets
set.seed(123) #uses a fixed seed so that every call will result in the same output.

USGS_cc4_lulc<-USGS_cc4_lulc[,c(36:38,29,26,2:16)]

intrain<-createDataPartition(y=USGS_cc4_lulc$detect ,
                             p=0.8,
                             list=FALSE)
#class(intrain) "matrix"

test<-USGS_cc4_lulc[-intrain,]
training<-USGS_cc4_lulc[intrain,]

dim(test)
dim(training)
```


<!-- #run the model -->
<!-- ```{r} -->

<!-- #Partitioning the data and running a GBM -->

<!-- #Creating training and test sets -->
<!-- set.seed(21) #uses a fixed seed so that every call will result in the same output. -->
<!-- intrain<-createDataPartition(y=USGS_cc4_lu21$detect , -->
<!--                              p=0.8, -->
<!--                              list=FALSE) -->
<!-- #class(intrain) "matrix" -->

<!-- test<-USGS_cc4_lu21[-intrain,] -->
<!-- training<-USGS_cc4_lu21[intrain,] -->

<!-- ``` -->

<!-- #fake big model -->
<!-- ```{r} -->
<!-- load("/Users/owner/Desktop/crypto/USGS_cc4_lu21.Rdata") -->
<!-- #USGS_cc4_lu21[,24]<-as.numeric(USGS_cc4_lu21[,24]) -->
<!-- fake<-USGS_cc4_lu21[,c(11,14,21,24)] -->
<!-- fake$DrainageAreaMeasure.MeasureValue2<-(runif(125, min= min(fake$DrainageAreaMeasure.MeasureValue) , max=max(fake$DrainageAreaMeasure.MeasureValue)) ) -->
<!-- fake$runoff2 <-(runif(125, min= min(fake$runoff ) , max=max(fake$runoff )) ) -->
<!-- fake$percent2<-(runif(125, min= 0 , max= 1) ) -->
<!-- head(fake) -->
<!-- set.seed(21) #uses a fixed seed so that every call will result in the same output. -->
<!-- intrain<-createDataPartition(y=fake$detect , -->
<!--                              p=0.8, -->
<!--                              list=FALSE) -->
<!-- #class(intrain) "matrix" -->

<!-- test<-fake[-intrain,] -->
<!-- training<-fake[intrain,] -->
<!-- model<-as.formula(paste(colnames(fake)[1], "~",      paste(colnames(fake)[c(2:7)], collapse="+"), #families -->
<!--                         sep = "")) -->
<!-- #that error message typically means you don’t have values for a variable represented in the model building process.  -->
<!-- water_gbm<- gbm(model, -->
<!--              data=training,  -->
<!--              distribution="bernoulli", -->
<!--              n.trees=100, -->
<!--              shrinkage=0.1, -->
<!--              interaction.depth=1, -->
<!--              bag.fraction=0.50, -->
<!--              #train.fraction=1, -->
<!--              n.minobsinnode=10, -->
<!--              cv.folds=10, -->
<!--              keep.data=TRUE, -->
<!--              class.stratify.cv=TRUE, -->
<!--              verbose="CV") -->

<!-- ``` -->

<!-- #the model -->
<!-- ```{r} -->
<!-- # generalized boosted regression model formula -->
<!-- #model<-as.formula(fake_glm_data$cry_detect~fake_glm_data$runoff+fake_glm_data$summer+fake_glm_data$watershed_size+fake_glm_data$urb+fake_glm_data$temp ) -->
<!-- USGS_cc4_lu21[,24]<-as.numeric(USGS_cc4_lu21[,24]) -->
<!-- model<-as.formula(paste(colnames(USGS_cc4_lu21)[11], "~",      paste(colnames(USGS_cc4_lu21)[c(14,21,24)], collapse="+"), #families -->
<!--                         sep = "")) -->
<!-- #that error message typically means you don’t have values for a variable represented in the model building process.  -->
<!-- water_gbm<- gbm(model, -->
<!--              data=training,  -->
<!--              distribution="bernoulli", -->
<!--              n.trees=100, -->
<!--              shrinkage=0.1, -->
<!--              interaction.depth=1, -->
<!--              bag.fraction=0.50, -->
<!--              #train.fraction=1, -->
<!--              n.minobsinnode=10, -->
<!--              cv.folds=10, -->
<!--              keep.data=TRUE, -->
<!--              class.stratify.cv=TRUE, -->
<!--              verbose="CV") -->

<!-- ``` -->

#look at the data
```{r}
boxplot(as.numeric(USGS_cc4_lulc$detect)~(USGS_cc4_lulc$GAGE_ID))
boxplot(log(as.numeric(USGS_cc4_lulc$runoff ))~(USGS_cc4_lulc$detect ))
boxplot((as.numeric(USGS_cc4_lulc$lulc21 ))~(USGS_cc4_lulc$detect ))
boxplot((as.numeric(USGS_cc4_lulc$lulc41 ))~(USGS_cc4_lulc$detect ))
boxplot((as.numeric(USGS_cc4_lulc$lulc81 ))~(USGS_cc4_lulc$detect ))
boxplot(log(as.numeric(USGS_cc4_lulc$DrainageAreaMeasure.MeasureValue ))~(USGS_cc4_lulc$detect ))
```

<!-- #as.factor -->
<!-- ```{r} -->
<!-- for (c in c(5,9,13,17,21,25,29,33,46,53)) -->
<!-- { -->
<!--  print (test[,c]%in%training[,c]) -->
<!--   print (class(test[,c])) -->
<!--   training[,c]<-as.factor(training[,c]) -->
<!--   test[,c]<-as.factor(test[,c]) -->
<!-- } -->
<!-- #test -->
<!-- #training -->

<!-- #43], "~",      paste(colnames(USGS_cc4_lulc)[c(5,9,13,17,21,25,29,33,46,53 -->
<!-- ``` -->


#the model
```{r}
# generalized boosted regression model formula
#model<-as.formula(fake_glm_data$cry_detect~fake_glm_data$runoff+fake_glm_data$summer+fake_glm_data$watershed_size+fake_glm_data$urb+fake_glm_data$temp )
#USGS_cc4_lu21[,24]<-as.numeric(USGS_cc4_lu21[,24])
# summary(lm(as.numeric(USGS_cc4_lulc$detect))~
#              USGS_cc4_lulc$lulc41 + 
#              USGS_cc4_lulc$lulc81 + 
#              USGS_cc4_lulc$lulc90 + 
#              USGS_cc4_lulc$lulc21 + 
#              USGS_cc4_lulc$lulc11 + 
#              USGS_cc4_lulc$lulc43 +
#              USGS_cc4_lulc$lulc52 +
#              USGS_cc4_lulc$lulc22 +
#              USGS_cc4_lulc$lulc23 +
#              USGS_cc4_lulc$DrainageAreaMeasure.MeasureValue + 
#              USGS_cc4_lulc$runoff+USGS_cc4_lulc$storm
#             # USGS_cc4_lulc$summer+USGS_cc4_lulc$winter
#            )
#USGS_cc4_lulc[,9]<-as.factor(USGS_cc4_lulc[,9])

model<-as.formula(paste(colnames(USGS_cc4_lulc)[5], "~",      paste(colnames(USGS_cc4_lulc)[c(1:4,6:20)], collapse="+"), #families
                        sep = ""))
#that error message typically means you don’t have values for a variable represented in the model building process. 
water_gbm<- gbm(model,
             data=training, 
             distribution="bernoulli",
             n.trees=400,
             shrinkage=0.1,
             interaction.depth=1,
             bag.fraction=0.50,
             #train.fraction=1,
             n.minobsinnode=10,
             cv.folds=10, # that error occurs when you have variable levels that exist in the test set which don't exist in the training set.
             keep.data=TRUE,
             class.stratify.cv=TRUE,
             verbose="CV")
gbm.perf(water_gbm)

```


<!-- #the model (clean input data) -->
<!-- ```{r} -->
<!-- # generalized boosted regression model formula -->
<!-- #model<-as.formula(fake_glm_data$cry_detect~fake_glm_data$runoff+fake_glm_data$summer+fake_glm_data$watershed_size+fake_glm_data$urb+fake_glm_data$temp ) -->
<!-- #USGS_cc4_lu21[,24]<-as.numeric(USGS_cc4_lu21[,24]) -->
<!-- #summary(lm(as.numeric(USGS_cc4_lulc$detect)~USGS_cc4_lulc$lulc41 + USGS_cc4_lulc$lulc81 + USGS_cc4_lulc$lulc90 + USGS_cc4_lulc$lulc21 + USGS_cc4_lulc$lulc11 + USGS_cc4_lulc$lulc43 +USGS_cc4_lulc$DrainageAreaMeasure.MeasureValue + USGS_cc4_lulc$runoff)) -->
<!-- #USGS_cc4_lulc<-USGS_cc4_lulc[,c(5,9,13,17,21,25,29,33,43,46,53)] -->
<!-- model<-as.formula(paste(colnames(USGS_cc4_lulc)[9], "~",      paste(colnames(USGS_cc4_lulc)[c(1:8,10,11 -->
<!--                                                                                                )], collapse="+"), #families -->
<!--                         sep = "")) -->
<!-- #that error message typically means you don’t have values for a variable represented in the model building process. -->
<!-- water_gbm<- gbm(model, -->
<!--              data=training, -->
<!--              distribution="bernoulli", -->
<!--              n.trees=100, -->
<!--              shrinkage=0.1, -->
<!--              interaction.depth=1, -->
<!--              bag.fraction=0.50, -->
<!--              #train.fraction=1, -->
<!--              n.minobsinnode=10, -->
<!--              cv.folds=10, # that error occurs when you have variable levels that exist in the test set which don't exist in the training set. -->
<!--              keep.data=TRUE, -->
<!--              class.stratify.cv=TRUE, -->
<!--              verbose="CV") -->

<!-- ``` -->


<!-- ```{r} -->
<!-- for (c in c(5,9,13,17,21,25,29,33,43,46,53)) -->
<!-- { -->
<!--  print (test[,c]%in%training[,c]) -->
<!--   print (class(test[,c])) -->
<!-- } -->
<!-- #test -->
<!-- #training -->

<!-- #43], "~",      paste(colnames(USGS_cc4_lulc)[c(5,9,13,17,21,25,29,33,46,53 -->
<!-- ``` -->


#lulc41 forest, #lulc90 wetland #21 developed #52	Shrub #11	Open Water #81	Pasture
```{r}
summary(water_gbm)
#boxplot(USGS_cc4_lulc$lulc90~USGS_cc4_lulc$detect )
#boxplot(log(as.numeric(USGS_cc4_lulc$runoff)) ~USGS_cc4_lulc$detect )

#percent positive in summer & winter
length(USGS_cc4_lulc[USGS_cc4_lulc$detect==1&USGS_cc4_lulc$summer==1,1])/length(USGS_cc4_lulc[USGS_cc4_lulc$summer==1,1])
length(USGS_cc4_lulc[USGS_cc4_lulc$detect==1&USGS_cc4_lulc$winter==1,1])/length(USGS_cc4_lulc[USGS_cc4_lulc$winter==1,1])


```

#predict on training
```{r}
gbmTrainPredictions = predict(object =water_gbm  ,
                              #newdata = train,
                             # n.trees = 100,
                              type = "response")
calibrate.plot(training$detect , gbmTrainPredictions, xlim=c(0,1), ylim=c(0,1))
```

#predict on test
```{r}
gbmTrainPredictions = predict(object =water_gbm  ,
                              newdata = test,
                             # n.trees = 100,
                              type = "response")

plot(test$detect,gbmTrainPredictions,ylim=c(0,1))
```


#run the model (only run mean LULC>10% (lulc21,22,41,81,82))
```{r}

USGS_cc4_lulc<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="TAVG"] <- "Temperature"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="storm"] <- "Surface_runoff"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc21"] <- "Developed_OpenSpace"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc22"] <- "Developed_LowIntensity"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc41"] <- "Forest"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc81"] <- "Pasture"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc82"] <- "Cultivated_Crops"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="detect"] <- "Detect"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="DrainageAreaMeasure.MeasureValue"] <- "Watershed_Area"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="runoff"] <- "Runoff"

#Partitioning the data and running a GBM

#Creating training and test sets
set.seed(123) #uses a fixed seed so that every call will result in the same output.

USGS_cc4_lulc<-USGS_cc4_lulc[,c(36:38,29,26,3:7)]

intrain<-createDataPartition(y=USGS_cc4_lulc$Detect ,
                             p=0.8,
                             list=FALSE)
#class(intrain) "matrix"

test<-USGS_cc4_lulc[-intrain,]
training<-USGS_cc4_lulc[intrain,]

dim(test)
dim(training)
```
#plot hist
```{r}
for (v in c(1:length(colnames(USGS_cc4_lulc))))
            {
  if (!class(USGS_cc4_lulc[,v])=="character")
  {
    jpeg(paste("/Users/owner/Desktop/crypto/cry/",colnames(USGS_cc4_lulc)[v],'.jpg',sep=''))

  hist(USGS_cc4_lulc[,v],xlab=colnames(USGS_cc4_lulc)[v],main="",ylim=c(0,100))
dev.off()
  }     
      }
```


#the model
```{r}
# generalized boosted regression model formula
#model<-as.formula(fake_glm_data$cry_detect~fake_glm_data$runoff+fake_glm_data$summer+fake_glm_data$watershed_size+fake_glm_data$urb+fake_glm_data$temp )
#USGS_cc4_lu21[,24]<-as.numeric(USGS_cc4_lu21[,24])

#USGS_cc4_lulc[,9]<-as.factor(USGS_cc4_lulc[,9])

model<-as.formula(paste(colnames(USGS_cc4_lulc)[5], "~",      paste(colnames(USGS_cc4_lulc)[c(1:4,6:10)], collapse="+"), #families
                        sep = ""))
#that error message typically means you don’t have values for a variable represented in the model building process. 
water_gbm<- gbm(model,
             data=training, 
             distribution="bernoulli",
             n.trees=600,
             shrinkage=0.1,
             interaction.depth=1,
             bag.fraction=0.50,
             #train.fraction=1,
             n.minobsinnode=10,
             cv.folds=10, # that error occurs when you have variable levels that exist in the test set which don't exist in the training set.
             keep.data=TRUE,
             class.stratify.cv=TRUE,
             verbose="CV")
# find index for n trees with minimum CV error
min_MSE <- which.min(water_gbm$cv.error)
print(min_MSE)
gbm.perf(water_gbm
         #,oobag.curve =TRUE
         ) #the optimal number of boosting iterations 
#an independent test set (test), out-of-bag estima- tion (OOB), and v-fold cross validation (cv)
#overlay a right y-axis is added to the training and test error plot and the estimated cumulative improvement in the loss function is plotted versus the iteration number.
#gbm.perf(water_gbm,method = "OOB")

#print the first and last trees
#plot(water_gbm$trees,water_gbm$train.error)
```

```{r}
summary(lm(USGS_cc4_lulc$Detect~USGS_cc4_lulc$Runoff+USGS_cc4_lulc$Surface_runoff+USGS_cc4_lulc$Temperature+USGS_cc4_lulc$Watershed_Area+USGS_cc4_lulc$Developed_OpenSpace+USGS_cc4_lulc$Developed_LowIntensity+USGS_cc4_lulc$Forest+USGS_cc4_lulc$Cultivated_Crops+USGS_cc4_lulc$Pasture))
```


#lulc41 forest, #lulc90 wetland #21 developed #52	Shrub #11	Open Water #81	Pasture
```{r}
par(mar = c(5, 11, 1, 1))
summary(
 water_gbm, 
  cBars = 10,
  method = relative.influence, # also can use permutation.test.gbm
  las = 2
  )
#boxplot(USGS_cc4_lulc$lulc90~USGS_cc4_lulc$detect )
boxplot(log(as.numeric(USGS_cc4_lulc$Surface_runoff )) ~USGS_cc4_lulc$Detect )

#percent positive in summer & winter
length(USGS_cc4_lulc[USGS_cc4_lulc$detect==1&USGS_cc4_lulc$summer==1,1])/length(USGS_cc4_lulc[USGS_cc4_lulc$summer==1,1])
length(USGS_cc4_lulc[USGS_cc4_lulc$detect==1&USGS_cc4_lulc$winter==1,1])/length(USGS_cc4_lulc[USGS_cc4_lulc$winter==1,1])
print(water_gbm)

```

#n a ROC curve the true positive rate (Sensitivity) is plotted in function of the false positive rate (100-Specificity) for different cut-off points of a parameter.
```{r}
preds <- predict.gbm( water_gbm, test,n.trees=
                        #water_gbm$gbm.call$best.trees
                        class(gbm.perf(water_gbm)),
                      type="response")
calc.deviance(obs=as.numeric(test$Detect ) , pred=preds
              , calc.mean=TRUE
              )
roc_obj <- roc(test$Detect ,preds)
auc(roc_obj)
plot(roc_obj)
plot(test$Detect ,preds)
predict_binary<-as.factor(ifelse(preds>0.5,1,0) )
test_obs<-as.factor(test$Detect)
confusionMatrix(predict_binary,test_obs)
```


#Partial dependence plots. After the most relevant variables have been identified, the next step is to attempt to understand how the response variable changes based on these variables. 
#PDPs plot the change in the average predicted value as specified feature(s) vary over their marginal distribution. 
```{r}
summary(water_gbm)
 water_gbm %>%
  partial(pred.var =  "Surface_runoff", n.trees = water_gbm$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)
#+scale_y_continuous(labels = scales::dollar)
 #+scale_y_continuous(labels = scales::marginal effect )

water_gbm %>%
  partial(pred.var =  "Forest", n.trees = water_gbm$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)
water_gbm %>%
  partial(pred.var =  "Cultivated_Crops", n.trees = water_gbm$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)
 water_gbm %>%
   partial(pred.var =  "Runoff", n.trees = water_gbm$n.trees, grid.resolution = 100) %>%
   autoplot(rug = TRUE, train = training)
water_gbm %>%
  partial(pred.var =  "Watershed_Area", n.trees = water_gbm$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)
water_gbm %>%
  partial(pred.var =  "Developed_OpenSpace", n.trees = water_gbm$n.trees, grid.resolution = 100) %>%autoplot(rug = TRUE, train = training)
water_gbm %>%
  partial(pred.var =  "Temperature", n.trees = water_gbm$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)
water_gbm %>%
  partial(pred.var =  "Pasture", n.trees = water_gbm$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)
```











#log
```{r}

USGS_cc4_lulc<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="TAVG"] <- "Temperature"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="storm"] <- "Surface_runoff"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc21"] <- "Developed_OpenSpace"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc22"] <- "Developed_LowIntensity"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc41"] <- "Forest"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc81"] <- "Pasture"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc82"] <- "Cultivated_Crops"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="detect"] <- "Detect"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="DrainageAreaMeasure.MeasureValue"] <- "Watershed_Area"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="runoff"] <- "log_runoff"

USGS_cc4_lulc_log<-USGS_cc4_lulc

USGS_cc4_lulc_log$log_runoff<-log(USGS_cc4_lulc_log$log_runoff)
#Partitioning the data and running a GBM

#Creating training and test sets
set.seed(123) #uses a fixed seed so that every call will result in the same output.

USGS_cc4_lulc_log<-USGS_cc4_lulc_log[,c(36:38,29,26,3:7)]

intrain<-createDataPartition(y=USGS_cc4_lulc_log$Detect ,
                             p=0.8,
                             list=FALSE)
#class(intrain) "matrix"

test<-USGS_cc4_lulc_log[-intrain,]
training<-USGS_cc4_lulc_log[intrain,]

dim(test)
dim(training)
```


#the model
```{r}
model_log<-as.formula(paste(colnames(USGS_cc4_lulc_log)[5], "~",      paste(colnames(USGS_cc4_lulc_log)[c(1:4,6:10)], collapse="+"), #families
                        sep = ""))
#that error message typically means you don’t have values for a variable represented in the model building process. 
water_gbm_log<- gbm(model_log,
             data=training, 
             distribution="bernoulli",
             n.trees=600,
             shrinkage=0.1,
             interaction.depth=1,
             bag.fraction=0.50,
             #train.fraction=1,
             n.minobsinnode=10,
             cv.folds=10, # that error occurs when you have variable levels that exist in the test set which don't exist in the training set.
             keep.data=TRUE,
             class.stratify.cv=TRUE,
             verbose="CV")

gbm.perf(water_gbm_log
         #,oobag.curve =TRUE
         ) #the optimal number of boosting iterations 
#an independent test set (test), out-of-bag estima- tion (OOB), and v-fold cross validation (cv)
#overlay a right y-axis is added to the training and test error plot and the estimated cumulative improvement in the loss function is plotted versus the iteration number.
#gbm.perf(water_gbm,method = "OOB")

#print the first and last trees
#plot(water_gbm$trees,water_gbm$train.error)
```


#lulc41 forest, #lulc90 wetland #21 developed #52	Shrub #11	Open Water #81	Pasture
```{r}
par(mar = c(5, 11, 1, 1))
summary(
 water_gbm_log, 
  cBars = 10,
  method = relative.influence, # also can use permutation.test.gbm
  las = 2
  )

print(water_gbm_log)

```




#find high runoff sampled watershed
```{r}
mean_runoff<-aggregate(USGS_cc4[,"runoff"], list(USGS_cc4$MonitoringLocationIdentifier ), mean)

median_runoff<-aggregate(USGS_cc4[,"runoff"], list(USGS_cc4$MonitoringLocationIdentifier ), median )

n_obs<-as.data.frame(table(USGS_cc4$ MonitoringLocationIdentifier))

mean_runoff_obs<-merge(mean_runoff,n_obs,by.x= "Group.1",by.y= "Var1")

mean_runoff_obs[mean_runoff_obs$x>3,]
mean_runoff_obs[mean_runoff_obs$x<1,]

```


#find high concentration sampled watershed (non detect sample w conc?)
```{r}
USGS_cc4$ResultMeasureValue<-as.numeric(USGS_cc4$ResultMeasureValue)
mean_conc<-aggregate(USGS_cc4[,"ResultMeasureValue" ], list(USGS_cc4$MonitoringLocationIdentifier), mean)

#median_runoff<-aggregate(USGS_cc4[,"runoff"], list(USGS_cc4$MonitoringLocationIdentifier ), median )

n_obs<-as.data.frame(table(USGS_cc4$ MonitoringLocationIdentifier))

mean_conc_obs<-merge(mean_conc,n_obs,by.x= "Group.1",by.y= "Var1")

mean_conc_obs[mean_conc_obs$x>60,]
mean_conc_obs[mean_conc_obs$x<30,]

USGS_04087214<-USGS_cc4[USGS_cc4$MonitoringLocationIdentifier=="USGS-04087214",]
par(mar=c(3,3,3,3))
plot(USGS_04087214$runoff ,USGS_04087214$ResultMeasureValue )
```

```{r}
par(mar=c(1,1,1,1))
plot(USGS_04087214$runoff ,USGS_04087214$ResultMeasureValue )
par(mar=c(1.5,1.5,1.5,1.5))
plot(USGS_04087214$month ,USGS_04087214$ResultMeasureValue )
```


#need to get Q in previous 48 days 
```{r}
USGS_cc4$storm<-" "
for (i in c(1:length(USGS_cc4$USGS)))
#for (i in c(41:59))
{
siteNo <- as.character(USGS_cc4[i,]$USGS)
pCode <- "00060" #discharge
start.date <- as.character(as.Date(USGS_cc4[i,]$ActivityStartDate,format=("%Y-%m-%d"))-48)
end.date <- as.character(USGS_cc4[i,]$ActivityStartDate)

if (nrow(readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = start.date,
                     endDate = end.date))>0)
{
  print(nrow(readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = start.date,
                     endDate = end.date)))
  USGS_cc4[i,]$Q<- readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = end.date,
                     endDate = end.date)$X_00060_00003
  
 st<- paste("Q_48days",i,sep='_')
  st<- readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = start.date,
                     endDate = end.date)
  print (paste("Q_48days",i,sep='_'))
  #hysep(s1372500$X105182_00060_00003,Dates=2/1/2019,da=1)
  #head(s1372500)
  #agency_cd site_no datetime X105182_00060_00003 X105182_00060_00003_cd
#1      USGS 1372500 2/1/2019                 399                      P
test<-hysep(as.integer(st$X_00060_00003),Dates=
              #2008-09-27
              #9/27/2008
              st$Date
            ,da=1)
print(test)
USGS_cc4[i,]$storm<-(1-test$BaseQ[49]/test$Flow[49])
}
else
{
 USGS_cc4[i,]$storm<-NA 
 USGS_cc4[i,]$Q<-NA 
}
  }
head(USGS_cc4)
plot(USGS_cc4$storm,as.numeric(USGS_cc4$runoff))
                                
```


#fix unit
```{r}
#summary(USGS_cc2$ResultMeasure.MeasureUnitCode)
               #/100L        #/l #/l                 % ocyst/100L oocyst/10L 
#       100          0          0          0          3         63        272  

#oocyst/10L
USGS_cc4$oocyst10L<-" "

USGS_cc4<-USGS_cc4[USGS_cc4$ResultMeasure.MeasureUnitCode=="oocyst/10L"|USGS_cc4$ResultMeasure.MeasureUnitCode=="ocyst/100L",]


summary(USGS_cc4$ResultMeasureValue)
i<-which(USGS_cc4$ResultMeasure.MeasureUnitCode=="oocyst/10L")
{
  USGS_cc4[i,]$oocyst10L<-as.numeric(as.character(USGS_cc4[i,]$ResultMeasureValue))
}

summary(as.numeric(USGS_cc4$oocyst10L))

j<-which(USGS_cc4$ResultMeasure.MeasureUnitCode=="ocyst/100L")
{
  USGS_cc4[j,]$oocyst10L<-as.numeric(as.character(USGS_cc4[j,]$ResultMeasureValue))/10
}
summary(as.numeric(USGS_cc4$oocyst10L))
```

```{r}
plot(USGS_cc4$runoff,USGS_cc4$oocyst10L)
plot(USGS_cc4$nlcd_2,USGS_cc4$oocyst10L)
plot(USGS_cc4$nlcd_8,USGS_cc4$oocyst10L)
plot(USGS_cc4$storm ,USGS_cc4$oocyst10L)
#plot(USGS_cc4$month ,USGS_cc4$oocyst10L)
boxplot(as.numeric(USGS_cc4$oocyst10L)~USGS_cc4$month,ylab="oocyst/10L",xlab="month")
boxplot(as.numeric(USGS_cc4$oocyst10L)~USGS_cc4$nlcd_2,ylab="oocyst/10L",xlab="month")
#plot(USGS_cc4$month,USGS_cc4$detect)
#plot(USGS_cc4[USGS_cc4$month=="6"|USGS_cc4$month=="7"|USGS_cc4$month=="8",]$runoff ,USGS_cc4[USGS_cc4$month=="6"|USGS_cc4$month=="7"|USGS_cc4$month=="8",]$oocyst10L)
plot(USGS_cc4$MonitoringLocationIdentifier,USGS_cc4$oocyst10L,ylab="oocyst/10L")
boxplot(as.numeric(USGS_cc4$oocyst10L)~USGS_cc4$MonitoringLocationIdentifier,ylab="oocyst/10L")
plot((USGS_cc4$DrainageAreaMeasure.MeasureValue),USGS_cc4$oocyst10L,log="x")
boxplot(as.numeric(USGS_cc4$oocyst10L)~USGS_cc4$DrainageAreaMeasure.MeasureValue ,ylab="oocyst/10L")
```

```{r}
USGS_cc4$oocyst10L<-as.numeric(USGS_cc4$oocyst10L)
USGS_cc4$runoff <-as.numeric(USGS_cc4$runoff )
USGS_cc4$DrainageAreaMeasure.MeasureValue<-as.numeric(USGS_cc4$DrainageAreaMeasure.MeasureValue)
USGS_cc4$storm<-as.numeric(USGS_cc4$storm)
USGS_cc4$nlcd_2_8<-as.numeric(USGS_cc4$nlcd_2+USGS_cc4$nlcd_8)
summary(lm(USGS_cc4$oocyst10L~USGS_cc4$runoff+USGS_cc4$DrainageAreaMeasure.MeasureValue+USGS_cc4$storm+USGS_cc4$nlcd_8))
```

```{r}
small<-USGS_cc4[USGS_cc4$DrainageAreaMeasure.MeasureValue<=25,]
plot(small$runoff,small$oocyst10L)
boxplot(as.numeric(small$oocyst10L)~small$MonitoringLocationIdentifier)
```


```{r}
write.csv(USGS_cc4,file="USGS_cc4.csv")
```


```{r}
USGS_cc2<-USGS_cc2[is.na(USGS_cc2$Q)==F,]

as.data.frame(table(USGS_cc2$MonitoringLocationIdentifier))
summary(USGS_cc2$ResultMeasureValue)
plot(USGS_cc2$Q,USGS_cc2$ResultMeasureValue)

USGS_01553500<-USGS_cc2[USGS_cc2$MonitoringLocationIdentifier=="USGS-01553500",]
plot(USGS_01553500$Q,USGS_01553500$ResultMeasureValue)

#count[count$Freq>09,]

USGS_03036500<-USGS_cc2[USGS_cc2$MonitoringLocationIdentifier=="USGS-03036500",]
plot(USGS_03036500$Q,USGS_03036500$ResultMeasureValue)

```

