#Load packages
```{r}
#install.packages("gbm")
library('gbm') #generalized boosted models
library('caTools') #calculates AUC
#install.packages("ROCR")
library('ROCR') #makes ROC curves
#install.packages("dismo")
library('dismo')
library('caret')
library("dataRetrieval")
library("pdp")          # model visualization
library("ggplot2")      # model visualization
library("lime")         # model visualization
#install.packages("pROC")
library("pROC")
#install.packages("e1071", repos="http://R-Forge.R-project.org")
library("e1071")
library( "MASS" )    #    used to generate correlated variables
library("sp")
library("dplyr")
library("Hmisc")    #   used for graphing se bars
#install.packages("rnoaa")
library("rnoaa")
```

#Read water quality data, remove -9999 and NA
```{r}
cc<-read.csv("C:/Users/gis/Documents/TAO/ecoli_20200201/narrowresult-EC_PA_WI_OH.tsv",header=T,sep="\t")
dim(cc)


cc<-cc[,c("ActivityStartTime.Time"                    ,"ActivityStartDate","MonitoringLocationIdentifier"          ,"CharacteristicName"
          #,"ResultSampleFractionText"
          , "ResultMeasureValue", "ResultMeasure.MeasureUnitCode"
          #,"ResultCommentText","ResultLaboratoryCommentText"
          )]



cc$ResultMeasureValue<-as.numeric(as.character(cc$ResultMeasureValue))
head(cc)
summary(cc$ResultMeasureValue)
cc<-cc[!(cc$ResultMeasureValue==-9999.0| is.na(cc$ResultMeasureValue)),]

cc$Detect<-as.numeric(0)

cc[cc$ResultMeasureValue>235,]$Detect<-as.numeric(1)

summary(cc$ResultMeasureValue)
dim(cc)
save(cc,file="C:/Users/gis/Documents/TAO/ecoli_20200201/cc.Rdata")
```

#read stations
```{r}
cc_s<-read.csv("C:/Users/gis/Documents/TAO/ecoli_20200201/station-EC_WI_PA_OH.csv",header=T,sep=',')

head(cc_s)
dim(cc_s)
cc_s<-cc_s[,c("MonitoringLocationIdentifier","MonitoringLocationTypeName","DrainageAreaMeasure.MeasureValue","DrainageAreaMeasure.MeasureUnitCode", "ContributingDrainageAreaMeasure.MeasureValue","ContributingDrainageAreaMeasure.MeasureUnitCode", "LatitudeMeasure", "LongitudeMeasure")]
head(cc_s)

dim(cc_s)
save(cc_s,file="C:/Users/gis/Documents/TAO/ecoli_20200201/cc_s.Rdata")
```

#find USGS "river" stations
```{r}
USGS_cc<-cc[grepl("USGS", cc$MonitoringLocationIdentifier)==T,]   
dim(USGS_cc)
USGS_cc$USGS<-USGS_cc$MonitoringLocationIdentifier

USGS_cc$USGS<-gsub('USGS-','',as.character(USGS_cc$USGS))
USGS_cc3<-USGS_cc[nchar(USGS_cc$USGS)==8,]

USGS_cc2<-merge(USGS_cc3,cc_s,by.x="MonitoringLocationIdentifier",by.y="MonitoringLocationIdentifier")

USGS_cc2<-USGS_cc2[!USGS_cc2$MonitoringLocationTypeName %in% c("Stream: Canal","Stream: Ditch"),]

print(paste("USGS_cc",dim(USGS_cc)))

print(paste("USGS_cc3",dim(USGS_cc3)))

print(paste("USGS_cc2",dim(USGS_cc2)))

save(USGS_cc2,file="C:/Users/gis/Documents/TAO/ecoli_20200201/USGS_cc2_0.Rdata")
```

#load LULC data
```{r}
setwd("C:/Users/gis/Documents/TAO/crypto/bas_ws")
CntlPlains_cry_ws<-load("bas_nonref_CntlPlains_cry_ws/bas_nonref_CntlPlains_cry_ws.Rdata")
EastHghlnds_cry_ws<-load("bas_nonref_EastHghlnds_cry_ws/bas_nonref_EastHghlnds_cry_ws.Rdata")
NorthEast_cry_ws<-load("bas_nonref_NorthEast_cry_ws/bas_nonref_NorthEast_cry_ws.Rdata")
```

#merge the lulc
```{r}
USGS_cc2_lulc_CntlPlains<-merge(bas_nonref_CntlPlains_cry_ws,USGS_cc2,by.x="GAGE_ID", by.y="MonitoringLocationIdentifier")
USGS_cc2_lulc_EastHghlnds<-merge(bas_nonref_EastHghlnds_cry_ws,USGS_cc2,by.x="GAGE_ID", by.y="MonitoringLocationIdentifier" )
USGS_cc2_lulc_NorthEast<-merge(bas_nonref_NorthEast_cry_ws,USGS_cc2,by.x="GAGE_ID", by.y="MonitoringLocationIdentifier" )

USGS_cc2_lulc_CntlPlains_EastHghlnds_cry_ws<-rbind(USGS_cc2_lulc_CntlPlains,USGS_cc2_lulc_EastHghlnds)
USGS_cc2_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws<-rbind(USGS_cc2_lulc_NorthEast,USGS_cc2_lulc_CntlPlains_EastHghlnds_cry_ws)

dim(USGS_cc2_lulc_CntlPlains)
dim(USGS_cc2_lulc_CntlPlains_EastHghlnds_cry_ws)
dim(USGS_cc2_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws)

dim(USGS_cc2)
dim(USGS_cc2[!USGS_cc2$MonitoringLocationIdentifier %in% USGS_cc2_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$GAGE_ID ,] ) #many of them are water quality stations, not Q
dim(USGS_cc2_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws)

save(USGS_cc2_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws,file="C:/Users/gis/Documents/TAO/ecoli_20200201/USGS_cc2_lulc_all_ws.Rdata")

```

#get Q 
```{r}
USGS_cc2<-USGS_cc2_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws
start_time<-Sys.time()

USGS_cc2$Q<-" "

for (i in c(1:length(USGS_cc2$USGS)))
{
  
siteNo <- as.character(USGS_cc2[i,]$USGS)
pCode <- "00060" #discharge
start.date <- as.character(USGS_cc2[i,]$ActivityStartDate)
end.date <- as.character(USGS_cc2[i,]$ActivityStartDate)
if (nrow(readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = start.date,
                     endDate = end.date))>0)
{
  USGS_cc2[i,]$Q<- readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = start.date,
                     endDate = end.date)$X_00060_00003
  print(i)
}
else
{
 USGS_cc2[i,]$Q<-NA 
}
  }
USGS_cc2$Q<-as.numeric(USGS_cc2$Q)
head(USGS_cc2)
save(USGS_cc2,file="C:/Users/gis/Documents/TAO/ecoli_20200201/USGS_cc2.Rdata")
print(Sys.time())
```

#get runoff (negative Q aND RUNOFF!) (USGS-04194085)
```{r}

USGS_cc4<-USGS_cc2[is.na(USGS_cc2$Q)==F,]

USGS_cc4$runoff<-""
#ft s/sq mi
#1 mi = 5280 ft.
USGS_cc4$runoff<-(as.numeric(USGS_cc4$Q)*86400)/((USGS_cc4$DrainageAreaMeasure.MeasureValue)*5280*5280)*(0.3048*1000) #mm/d

USGS_cc4$log_Q<-log(USGS_cc4$Q)
# independent variable was more than two standard deviations away from the mean.
#Christensen, Victoria G., Xiaodong Jian, and Andrew C. Ziegler. Regression analysis and real-time water-quality monitoring to estimate constituent concentrations, loads, and yields in the Little Arkansas River, south-central Kansas, 1995-99. No. 2000-4126. US Department of the Interior, US Geological Survey; Information Services [distributor],, 2000.

#https://books.google.com.tw/books?id=OhkTXdpp77oC&pg=RA5-PA24&lpg=RA5-PA24&dq=remove+discharge+outlier+regression+water+quality&source=bl&ots=wojbXc436x&sig=ACfU3U0GndpSeSjsza-XojYfCfVb3N6vBQ&hl=en&sa=X&ved=2ahUKEwig4pak_qDnAhWKuJQKHcRDBhkQ6AEwAXoECAoQAQ#v=onepage&q=outliers&f=false
#2*sd(USGS_cc4$Q)+mean(USGS_cc4$Q)

USGS_cc4<-USGS_cc4[USGS_cc4$log_Q<(2*sd(USGS_cc4$log_Q)+mean(USGS_cc4$log_Q)) & USGS_cc4$log_Q>(-2*sd(USGS_cc4$log_Q)+mean(USGS_cc4$log_Q)),]

dim(USGS_cc4)

save(USGS_cc4,file="C:/Users/gis/Documents/TAO/ecoli_20200201/USGS_cc4_202002.Rdata")
```

#subset year to 2002~2009
```{r}


USGS_cc4$ActivityStartDate<-as.Date(as.character(USGS_cc4$ActivityStartDate),format = "%Y-%m-%d")
USGS_cc4<-USGS_cc4[USGS_cc4$ActivityStartDate>=as.Date("2002-01-01",format = "%Y-%m-%d") &  USGS_cc4$ActivityStartDate<=as.Date("2009-12-31",format = "%Y-%m-%d"),]
dim(USGS_cc4)
```

#find the nearest Tmax and Tmin climate station for each USGS station
```{r}
USGS_cc4_lulc<-USGS_cc4
m<-matrix(c(USGS_cc4_lulc$LongitudeMeasure,USGS_cc4_lulc$LatitudeMeasure),ncol=2)
mu<-unique(m)

load("C:/Users/gis/Documents/TAO/crypto/stations.Rdata")

station_last<-stations[stations$last_year>2009 & stations$first_year<2002 ,]
#temperature
station_last_TMAX_TMIN<-station_last[station_last$element %in% c("TMAX"),]
station_last_TMAX_TMIN<-station_last_TMAX_TMIN[station_last_TMAX_TMIN$state %in% c("OH","PA","WI"),]

sm<-matrix(c(station_last_TMAX_TMIN$longitude,station_last_TMAX_TMIN$latitude ),ncol=2)
smu<-unique(sm)

lon<-as.numeric("")
lat<-as.numeric("")
USGS_lon<-as.numeric("")
df<-data.frame(
  lon,lat,USGS_lon)

for (i in c(1:dim(mu)[1]))
{
  km <- spDistsN1(smu, #climate station
                  mu[i,], longlat=TRUE)
  i2<-which(km==min(km))
  sdf<-data.frame(smu[i2,1],smu[i2,2],mu[i,][1])
  colnames(sdf)<-c("lon", "lat","USGS_lon")
  df<-rbind(df,sdf)
}
summary(df) #list of weather stations (can be repeated)
```

#the table of NOAA & USGS stations
```{r}
dim(station_last_TMAX_TMIN)
station_last_TMAX_TMIN_nearest<-merge(station_last_TMAX_TMIN, #NOAA stations
                                      df[-1,], #USGS lon  and the nearest NOAA stations
                                      by.x="latitude",by.y="lat")

dim(station_last_TMAX_TMIN_nearest)


save(station_last_TMAX_TMIN_nearest,file="C:/Users/gis/Documents/TAO/ecoli_20200201/station_last_TMAX_TMIN_nearest.Rdata")
```

#1. (USGS_cc4_lulc) LULC-date-ResultMeasureValue-drainage area-"LatitudeMeasure"- "LongitudeMeasure"-"Q"-"runoff"
#2. (station_last_TMAX_TMIN_nearest) "latitude"-"id"-"longitude"-"elevation"-"state"-"name"-"gsn_flag"-"wmo_id"-"element"- "first_year" "last_year"-"lon"-"USGS_lon"
#3. merge by "USGS_lon"
```{r}

USGS_cc4_lulc_nearest<-merge(USGS_cc4_lulc,station_last_TMAX_TMIN_nearest,by.x= "LongitudeMeasure" ,by.y=  "USGS_lon")
dim(USGS_cc4_lulc_nearest)

save(USGS_cc4_lulc_nearest,file="C:/Users/gis/Documents/TAO/ecoli_20200201/USGS_cc4_lulc_nearest.Rdata")
# id - ResultMeasureValue - RUNOFF- DATE - DrainageAreaMeasure.MeasureValu- LULC
```

#get tmin 
```{r}
u_sta<-unique( USGS_cc4_lulc_nearest$id)

id<-c("")
 tmin<-as.numeric("")
 date<-as.Date("1234-12-21")     
 mflag<-c("") 
 qflag<-c("") 
 sflag<-c("")
 sta_df<-data.frame(id,tmin,date,mflag,qflag,sflag)
 sta_df$id<-as.character(sta_df$id)
 
 for (s in c(1:length(u_sta)))
{
  u_sta[s]
  g<-ghcnd_search(u_sta[s], var = "TMIN")
  sta_tmin<-g$tmin
   sta_df<-rbind( sta_df,  sta_tmin)
 }

colnames(sta_df)<-c("id","tmin" ,"ActivityStartDate","mflag","qflag","sflag")    
USGS_cc4_lulc_nearest$ActivityStartDate<-as.Date(USGS_cc4_lulc_nearest$ActivityStartDate) 
sta_df$id<-as.character(sta_df$id)
USGS_cc4_lulc_nearest_tmin<- full_join(USGS_cc4_lulc_nearest, sta_df,by = c("id", "ActivityStartDate"))
USGS_cc4_lulc_nearest_tmin_rmna<-USGS_cc4_lulc_nearest_tmin[!is.na(USGS_cc4_lulc_nearest_tmin$Detect  ),]
dim(USGS_cc4_lulc_nearest_tmin_rmna)
USGS_cc4_lulc_nearest_tmin_rmna<-USGS_cc4_lulc_nearest_tmin_rmna[USGS_cc4_lulc_nearest_tmin_rmna$element=="TMAX",]
dim(USGS_cc4_lulc_nearest_tmin_rmna)
save(USGS_cc4_lulc_nearest_tmin_rmna,file="C:/Users/gis/Documents/TAO/ecoli_20200201/USGS_cc4_lulc_nearest_tmin.Rdata")

```

#get tmax 
```{r}
id<-c("")
 tmax<-as.numeric("")
 date<-as.Date("1234-12-21")     
 mflag<-c("") 
 qflag<-c("") 
 sflag<-c("")
 sta_df_max<-data.frame(id,tmax,date,mflag,qflag,sflag)
 
 for (s in c(1:length(u_sta)))
{
  u_sta[s]
  g<-ghcnd_search(u_sta[s], var = "TMAX")
  sta_tmax<-g$tmax
   sta_df_max<-rbind( sta_df_max,  sta_tmax)
 }

colnames(sta_df_max)<-c("id","tmax" ,"ActivityStartDate","mflag","qflag","sflag")    
sta_df_max$id<-as.character(sta_df_max$id)

USGS_cc4_lulc_nearest_tmin_rmna_tmax<- full_join(USGS_cc4_lulc_nearest_tmin_rmna, sta_df_max,by = c("id", "ActivityStartDate"))
summary(USGS_cc4_lulc_nearest_tmin_rmna_tmax)

USGS_cc4_lulc_nearest_tmin_rmna_tmax<-USGS_cc4_lulc_nearest_tmin_rmna_tmax[!is.na(USGS_cc4_lulc_nearest_tmin_rmna_tmax$Detect) ,]

save(USGS_cc4_lulc_nearest_tmin_rmna_tmax,file="C:/Users/gis/Documents/TAO/ecoli_20200201/USGS_cc4_lulc_nearest_tmin_rmna_tmax.Rdata")

```

#find the stations with the most TAVG
```{r}
id<-c("")
 tavg<-as.numeric("")
 date<-as.Date("1234-12-21")     
 mflag<-c("") 
 qflag<-c("") 
 sflag<-c("")
 sta_df_avg2<-data.frame(id,tavg,date,mflag,qflag,sflag)

 load("/Users/owner/Desktop/crypto202001/stations.Rdata")
 stations<-stations[stations$state %in% c("OH","PA","WI","IL"),]
  stations<- stations[ stations$element=="TAVG",]
  stations<-stations[stations$last_year>2009 & stations$first_year<2002 ,]
  u_sta2<-unique(stations$id )

 for (s in c(1:length(u_sta2)))
{
  u_sta2[s]
  g<-ghcnd_search(u_sta2[s], var = "TAVG",date_min = "2002-01-01", date_max = "2009-12-31")
  sta_tavg2<-g$tavg
   sta_df_avg2<-rbind( sta_df_avg2,  sta_tavg2)
 }
 table(   sta_df_avg2$id)
```

```{r}
            USR0000PALL USR0000PERI USR0000WANT USR0000WAUG USR0000WBAR USR0000WBRF USR0000WDDG 
          1        1947        2922        2498        2593        2499        2737        2803 
USR0000WGLI USW00013739 USW00014711 USW00014737 USW00014777 USW00014778 USW00014820 USW00014821 
       2744        1308        1308        1308        1308        1308        1308        1308 
USW00014837 USW00014839 USW00014842 USW00014852 USW00014860 USW00014891 USW00014895 USW00014898 
       1308        1308        1308        1308        1308        1308        1308        1308 
USW00014920 USW00014923 USW00093815 USW00093822 USW00093989 USW00094822 USW00094823 USW00094830 
       1308        1308        1308        1308        1308        1308        1308        1308 
USW00094846 
       1308
       #>2000
       stations[stations$id%in% c("USR0000PERI", "USR0000WANT", "USR0000WAUG", "USR0000WBAR", "USR0000WBRF", "USR0000WDDG", "USR0000WGLI"),]
       
```

#find the nearest TAVG climate station for each USGS station
```{r}


m<-matrix(c(USGS_cc4_lulc$LongitudeMeasure,USGS_cc4_lulc$LatitudeMeasure),ncol=2)
mu<-unique(m)
load("C:/Users/gis/Documents/TAO/crypto/stations.Rdata")
stations<-stations[(stations$id%in% c("USR0000PERI", "USR0000WANT", "USR0000WAUG", "USR0000WBAR", "USR0000WBRF", "USR0000WDDG", "USR0000WGLI")),]
station_last<-stations[stations$last_year>2009 & stations$first_year<2002 ,]
#temperature
station_last_TAVG<-station_last[station_last$element %in% c("TAVG"),]
station_last_TAVG<-station_last_TAVG[station_last_TAVG$state %in% c("OH","PA","WI","IL"),]

sm<-matrix(c(station_last_TAVG$longitude,station_last_TAVG$latitude ),ncol=2)
smu<-unique(sm)

lon<-as.numeric("")
lat<-as.numeric("")
USGS_lon<-as.numeric("")
df<-data.frame(
  lon,lat,USGS_lon)

for (i in c(1:dim(mu)[1]))
{
  km <- spDistsN1(smu, #climate station
                  mu[i,], longlat=TRUE)
  i2<-which(km==min(km))
  sdf<-data.frame(smu[i2,1],smu[i2,2],mu[i,][1])
  colnames(sdf)<-c("lon", "lat","USGS_lon")
  df<-rbind(df,sdf)
}
summary(df) #list of weather stations (can be repeated)
```

#the table of NOAA & USGS stations
```{r}

station_last_TAVG_nearest<-merge(station_last_TAVG, #NOAA stations
                                      df[-1,], #USGS lon  and the nearest NOAA stations
                                      by.x="latitude",by.y="lat")


colnames(station_last_TAVG_nearest)<-c( colnames(station_last_TAVG_nearest)[1],"tavg_id", colnames(station_last_TAVG_nearest)[3:13])

save(station_last_TAVG_nearest,file="C:/Users/gis/Documents/TAO/ecoli_20200201/station_last_TAVG_nearest.Rdata")
```

#get TAVG
```{r}
id<-c("")
 tavg<-as.numeric("")
 date<-as.Date("1234-12-21")     
 mflag<-c("") 
 qflag<-c("") 
 sflag<-c("")
 sta_df_avg<-data.frame(id,tavg,date,mflag,qflag,sflag)
 
 u_sta<-unique(station_last_TAVG_nearest$tavg_id )
 
 for (s in c(1:length(u_sta)))
{
  u_sta[s]
  g<-ghcnd_search(u_sta[s], var = "TAVG",date_min = "2002-01-01", date_max = "2009-12-31")
  sta_tavg<-g$tavg
   sta_df_avg<-rbind( sta_df_avg,  sta_tavg)
 }

colnames(sta_df_avg)<-c("tavg_id","tavg" ,"ActivityStartDate","mflag","qflag","sflag")    
sta_df_avg$tavg_id<-as.character(sta_df_avg$tavg_id)
table(sta_df_avg$tavg_id)
USGS_cc4_lulc_nearest_tmin_rmna_tmax<-merge(USGS_cc4_lulc_nearest_tmin_rmna_tmax,station_last_TAVG_nearest,by.x="LongitudeMeasure",by.y="USGS_lon")
dim(USGS_cc4_lulc_nearest_tmin_rmna_tmax)
USGS_cc4_lulc_nearest_tmin_rmna_tmax_tavg<- full_join(USGS_cc4_lulc_nearest_tmin_rmna_tmax, sta_df_avg,by = c("tavg_id", "ActivityStartDate"))
summary(USGS_cc4_lulc_nearest_tmin_rmna_tmax_tavg)

USGS_cc4_lulc_nearest_tmin_rmna_tmax_tavg<-USGS_cc4_lulc_nearest_tmin_rmna_tmax_tavg[!is.na(USGS_cc4_lulc_nearest_tmin_rmna_tmax_tavg$Detect) ,]

dim(USGS_cc4_lulc_nearest_tmin_rmna_tmax_tavg)
#table(USGS_cc4_lulc_nearest_tmin_rmna_tmax_tavg[is.na(USGS_cc4_lulc_nearest_tmin_rmna_tmax_tavg$tavg) ,]$id)
save(USGS_cc4_lulc_nearest_tmin_rmna_tmax_tavg,file="C:/Users/gis/Documents/TAO/ecoli_20200201/USGS_cc4_lulc_nearest_tmin_rmna_tmax_tavg.Rdata")

```



#get PRCP and TAVG

#hysep function
```{r}

hysep <- function(Flow, Dates, Start=NULL, End=NULL, da,
                 select="sliding", STAID="Unknown") {
  ## Start of code: initial processing
  STAID <- as.character(STAID[1L])
  if(is.null(Start))
    Start <- Dates[1L]
  else if(is.character(Start))
    Start <- as.Date(Start)
  if(is.null(End))
    End <- Dates[length(Dates)]
  else if(is.character(End))
    End <- as.Date(End)
  sel <- (Dates >= Start) & (Dates <= End)
  Dates <- Dates[sel]
  Flow <- pmax(Flow[sel], 1e-99) # Convert 0 to a small number
  if(any(is.na(Flow)))
    stop("Missing values between ", Start, " and ", End)
  if(any(diff(as.double(Dates)) != 1))
    stop("Date data are not continuous between Start and End")
  select <- match.arg(select, c("sliding", "local minimum", "fixed"))
  Nact <- max(da^0.2, 1)
  N2star <- max((((2*Nact) ) %/% 2)*2 + 1, 3)
  ## Set up for fixed--construct intervals of length N2star
  Nobs <- length(Flow)
  Ngrp <- ceiling(Nobs / N2star)
  Grps <- inverse.rle(list(lengths=rep(N2star, Ngrp), values=seq(Ngrp)))
  length(Grps) <- Nobs # Truncate if necessary
  ## Compute the fixed method
  Mins <- tapply(Flow, Grps, min)
  Fixed <- Mins[as.character(Grps)]
  ## Now the sliding method
  Slide <- sapply(seq(N2star, Nobs), function(i)
    min(Flow[seq(i - N2star + 1L, i)])
  )
  SlB <- Slide[1L]
  SlE <- Slide[length(Slide)]
  Nfil <- (N2star - 1L) / 2L
  Slide <- c(rep(SlB, Nfil), Slide, rep(SlE, Nfil))
  ## And the local minimum
  Mid <- as.integer((N2star) / 2)
  LocMin <- sapply(seq(N2star, Nobs), function(i)
    min(Flow[seq(i - N2star + 1L, i)]) == Flow[i - Mid]
  )
  LocMin <- c(rep(FALSE, Nfil), LocMin, rep(FALSE, Nfil))
  ## Need to trap short periods where only 1 local minimum
  if(sum(LocMin) == 1L) {
    warning("Only one local minimum in calibration period")
    LocMin <- pmax(Flow[LocMin], 0.01)
  } else
    LocMin <- exp(approx(which(LocMin), log(pmax(Flow[LocMin], 0.01)), xout=seq(Nobs), rule=2)$y)
  LocMin <- pmin(Flow, LocMin) # recover 0s and tails
  if(select == "fixed")
    BaseQ <- Fixed
  else if(select == "sliding")
    BaseQ <- Slide
  else 
    BaseQ <- LocMin
  retval <- data.frame(Dates=Dates, BaseQ=round(BaseQ, 3L), 
                       Flow=Flow, Fixed=Fixed, Sliding=Slide, 
                       LocalMin=LocMin)
  if(!is.null(STAID))
    attr(retval, "STAID") <- STAID
  attr(retval, "type") <- "hysep"
  class(retval) <- c("baseflow", "data.frame")
  return(retval)
}

```

#need to get Q in previous 44 days 
```{r}
USGS_cc4<- USGS_cc4_lulc_nearest_tmin_rmna_tmax_tavg[!(is.na(USGS_cc4_lulc_nearest_tmin_rmna_tmax_tavg$tavg) | is.na(USGS_cc4_lulc_nearest_tmin_rmna_tmax_tavg$tmax)  | is.na(USGS_cc4_lulc_nearest_tmin_rmna_tmax_tavg$tmin ) ) ,] 
print(Sys.time())
USGS_cc4$storm<-" "
for (i in c(1:length(USGS_cc4$USGS)))
{
siteNo <- as.character(USGS_cc4[i,]$USGS)
pCode <- "00060" #discharge
start.date <- as.character(as.Date(USGS_cc4[i,]$ActivityStartDate,format=("%Y-%m-%d"))-44)
end.date <- as.character(USGS_cc4[i,]$ActivityStartDate)

if (nrow(readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = start.date,
                     endDate = end.date))>0)
{
  print(nrow(readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = start.date,
                     endDate = end.date)))
  USGS_cc4[i,]$Q<- readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = end.date,
                     endDate = end.date)$X_00060_00003
  
 st<- paste("Q_44days",i,sep='_')
  st<- readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = start.date,
                     endDate = end.date)
  print (paste("Q_44days",i,sep='_'))

  
test<-hysep(as.integer(st$X_00060_00003),Dates=
              #2008-09-27
              #9/27/2008
              st$Date
            ,da=1)
print(test)
USGS_cc4[i,]$storm<-(1-test$BaseQ[45]/test$Flow[45])
}
else
{
 USGS_cc4[i,]$storm<-NA 
 USGS_cc4[i,]$Q<-NA 
}
print(paste("day ",i," done",sep=','))
  }
head(USGS_cc4)
plot(USGS_cc4$storm,as.numeric(USGS_cc4$runoff))
print(Sys.time())
save(USGS_cc4,file="C:/Users/gis/Documents/TAO/ecoli_20200201/USGS_cc4_20190507_2.Rdata")
```

```{r}
load("/Users/owner/Desktop/crypto/ec/USGS_cc4_20190507_2.Rdata")
#usgs_cc4 (USGS_lon_date)
USGS_cc4$storm<-as.numeric(USGS_cc4$storm)
summary(USGS_cc4$storm)
USGS_cc4<-USGS_cc4[!is.na(USGS_cc4$storm),]
USGS_cc4<-USGS_cc4[!is.na(USGS_cc4$tmax),]
USGS_cc4$Q<-as.numeric(USGS_cc4$Q)
summary(USGS_cc4)
#"storm", "tmax"
```


#binary detect
```{r}
USGS_cc4_lulc<-USGS_cc4
dim(USGS_cc4_lulc)

# USGS_cc4_lulc$Detect<-as.numeric(0)
# USGS_cc4_lulc[USGS_cc4$ResultMeasureValue>235,]$Detect<-as.numeric(1)
# dim(USGS_cc4_lulc)

USGS_cc4_lulc$Developed<-USGS_cc4_lulc$lulc21    + USGS_cc4_lulc$lulc22+USGS_cc4_lulc$lulc23+USGS_cc4_lulc$lulc24
USGS_cc4_lulc$Forest <-USGS_cc4_lulc$lulc41    + USGS_cc4_lulc$lulc42 +USGS_cc4_lulc$lulc43
USGS_cc4_lulc$Other<-USGS_cc4_lulc$lulc11 +USGS_cc4_lulc$lulc31 +USGS_cc4_lulc$lulc52  +USGS_cc4_lulc$lulc71 +USGS_cc4_lulc$lulc90 + USGS_cc4_lulc$lulc95 #water #Barren # Shrubland #Herbaceous #wetland  
USGS_cc4_lulc$Cultivated<-USGS_cc4_lulc$lulc81 +	 USGS_cc4_lulc$lulc82

USGS_cc4_lulc<-USGS_cc4_lulc[!is.na(USGS_cc4_lulc$tmin),]
USGS_cc4_lulc$tmin<-USGS_cc4_lulc$tmin/10
USGS_cc4_lulc$tmax<-USGS_cc4_lulc$tmax/10

USGS_cc4_lulc$Forest_Min_temperature<-USGS_cc4_lulc$Forest*USGS_cc4_lulc$tmin
USGS_cc4_lulc$Forest_Max_temperature <-USGS_cc4_lulc$Forest*USGS_cc4_lulc$tmax
USGS_cc4_lulc$Forest_Surface_runoff <- USGS_cc4_lulc$Forest*USGS_cc4_lulc$storm
USGS_cc4_lulc$Forest_Runoff <- USGS_cc4_lulc$Forest*USGS_cc4_lulc$runoff

USGS_cc4_lulc$Developed_Min_temperature<-USGS_cc4_lulc$Developed*USGS_cc4_lulc$tmin
USGS_cc4_lulc$Developed_Max_temperature <-USGS_cc4_lulc$Developed*USGS_cc4_lulc$tmax
USGS_cc4_lulc$Developed_Surface_runoff <- USGS_cc4_lulc$Developed*USGS_cc4_lulc$storm
USGS_cc4_lulc$Developed_Runoff <- USGS_cc4_lulc$Developed*USGS_cc4_lulc$runoff

USGS_cc4_lulc$Other_Min_temperature<-USGS_cc4_lulc$Other*USGS_cc4_lulc$tmin
USGS_cc4_lulc$Other_Max_temperature <-USGS_cc4_lulc$Other*USGS_cc4_lulc$tmax
USGS_cc4_lulc$Other_Surface_runoff <- USGS_cc4_lulc$Other*USGS_cc4_lulc$storm
USGS_cc4_lulc$Other_Runoff <- USGS_cc4_lulc$Other*USGS_cc4_lulc$runoff

USGS_cc4_lulc$Cultivated_Min_temperature<-USGS_cc4_lulc$Cultivated*USGS_cc4_lulc$tmin
USGS_cc4_lulc$Cultivated_Max_temperature <-USGS_cc4_lulc$Cultivated*USGS_cc4_lulc$tmax
USGS_cc4_lulc$Cultivated_Surface_runoff <- USGS_cc4_lulc$Cultivated*USGS_cc4_lulc$storm
USGS_cc4_lulc$Cultivated_Runoff <- USGS_cc4_lulc$Cultivated*USGS_cc4_lulc$runoff

#USGS_cc4_lulc$Surface_runoff_Min_temperature <-USGS_cc4_lulc$storm*USGS_cc4_lulc$tmin
#USGS_cc4_lulc$Surface_runoff_Max_temperature <-USGS_cc4_lulc$storm*USGS_cc4_lulc$tmax
#USGS_cc4_lulc$Surface_runoff_Runoff <-USGS_cc4_lulc$storm*USGS_cc4_lulc$runoff

colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="storm"] <-"Surface_runoff"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="tmin"] <- "Min_temperature"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="tmax"] <- "Max_temperature"

dim(USGS_cc4_lulc)
USGS_cc4_lulc$Tavg<-(USGS_cc4_lulc$Max_temperature+USGS_cc4_lulc$Min_temperature)/2

save(USGS_cc4_lulc,file="/Users/owner/Desktop/crypto/ec/USGS_cc4_201912.Rdata")

```

```{r}
USGS_cc5<-USGS_cc4_lulc

summary(USGS_cc5$ocyst100)

#extract month

USGS_cc5$mo<-months(USGS_cc5$ActivityStartDate )
USGS_cc5_by_month<-aggregate(USGS_cc5$Detect ,by=list(c(USGS_cc5$mo )),FUN="sum")
table(USGS_cc5$mo)
mo<-as.data.frame(table(USGS_cc5$mo))
USGS_cc5_by_month<-merge(USGS_cc5_by_month,mo,by.x="Group.1",by.y="Var1")
month_df <-as.data.frame(month.name)

month_nu<-as.data.frame((c(4,8,12,2,1,7,6,3,5,11,10,9)))

USGS_cc5_by_month$perc<-USGS_cc5_by_month$x/USGS_cc5_by_month$Freq
USGS_cc5_by_month<-merge(month_df,USGS_cc5_by_month,by.x="month.name",by.y="Group.1",all=T)
USGS_cc5_by_month<-cbind(USGS_cc5_by_month,month_nu)
USGS_cc5_by_month<-USGS_cc5_by_month[order(USGS_cc5_by_month$`(c(4, 8, 12, 2, 1, 7, 6, 3, 5, 11, 10, 9))`),]
colnames(USGS_cc5_by_month)<-c("month.name","x","Freq","Proportion","Month")

jpeg("/Users/owner/Desktop/crypto/ec/month_detect.jpg")
plot(USGS_cc5_by_month$Month,USGS_cc5_by_month$Proportion ,xlab = "Month",ylab = "Proportion of E. coli +",ylim=c(0.4,1))
dev.off()

```

#split the data
```{r}
load("/Users/owner/Desktop/crypto/ec/USGS_cc4_201912.Rdata")
#Partitioning the data and running a GBM

#Creating training and test sets
set.seed(123) #uses a fixed seed so that every call will result in the same output.

USGS_cc4_lulc<-USGS_cc4_lulc[,c(52:55  #LULC
                                ,42,46 #TEMPERATURE
                                ,50 #SURFACE Q
                                ,29 #RUNOFF
                                ,23 #WATERSHED AREA
                                ,51 #DETECT
                                ,56:71 #interactive
                                )]

intrain<-createDataPartition(y=USGS_cc4_lulc$Detect ,
                             p=0.8,
                             list=FALSE)
#class(intrain) "matrix"

test<-USGS_cc4_lulc[-intrain,]
training<-USGS_cc4_lulc[intrain,]

dim(test)
dim(training)
```

#split the data (Tavg)
```{r}
load("/Users/owner/Desktop/crypto/ec/USGS_cc4_201912.Rdata")
#Partitioning the data and running a GBM

#Creating training and test sets
set.seed(123) #uses a fixed seed so that every call will result in the same output.

USGS_cc4_lulc<-USGS_cc4_lulc[,c(52:55  #LULC
                                ,72 #TEMPERATURE
                                ,50 #SURFACE Q
                                ,29 #RUNOFF
                                ,23 #WATERSHED AREA
                                ,51 #DETECT
                                ,56:71 #interactive
                                )]

intrain<-createDataPartition(y=USGS_cc4_lulc$Detect ,
                             p=0.8,
                             list=FALSE)
#class(intrain) "matrix"

test<-USGS_cc4_lulc[-intrain,]
training<-USGS_cc4_lulc[intrain,]

dim(test)
dim(training)
save(test,file="/Users/owner/Desktop/crypto/ec/USGS_cc4_test.Rdata")
```

#the model
```{r}
model<-as.formula(paste(colnames(USGS_cc4_lulc)[10], "~",      paste(colnames(USGS_cc4_lulc)[c(1:9,11:26)], collapse="+"), #families
                        sep = ""))
#that error message typically means you don’t have values for a variable represented in the model building process. 
water_gbm_lumped<- gbm(model,
             data=training, 
             distribution="bernoulli",
             n.trees=1900,
             shrinkage=0.1,
             interaction.depth=1,
             bag.fraction=0.50,
             #train.fraction=1,
             n.minobsinnode=10,
            cv.folds=10, # that error occurs when you have variable levels that exist in the test set which don't exist in the training set.
             keep.data=TRUE,
             class.stratify.cv=TRUE,
             verbose="CV")
#jpeg(paste("/Users/owner/Desktop/crypto/ec/","lumped_perf",'.jpg',sep=''))
gbm.perf(water_gbm_lumped)
save(water_gbm_lumped,file="/Users/owner/Desktop/crypto/ec/water_gbm_lumped.Rdata")
```

#the simple model
```{r}
model<-as.formula(paste(colnames(USGS_cc4_lulc)[10], "~",      paste(colnames(USGS_cc4_lulc)[c(1:9)], collapse="+"), #families
                        sep = ""))
#that error message typically means you don’t have values for a variable represented in the model building process. 
water_gbm_lumped<- gbm(model,
             data=training, 
             distribution="bernoulli",
             n.trees=1200,
             shrinkage=0.1,
             interaction.depth=1,
             bag.fraction=0.50,
             #train.fraction=1,
             n.minobsinnode=10,
            cv.folds=10, # that error occurs when you have variable levels that exist in the test set which don't exist in the training set.
             keep.data=TRUE,
             class.stratify.cv=TRUE,
             verbose="CV")
#jpeg(paste("/Users/owner/Desktop/crypto/ec/","lumped_perf",'.jpg',sep=''))
gbm.perf(water_gbm_lumped)
save(water_gbm_lumped,file="/Users/owner/Desktop/crypto/ec/water_gbm_simple.Rdata")
```


#the simple model (Tavg)
```{r}

model<-as.formula(paste(colnames(USGS_cc4_lulc)[9], "~",      paste(colnames(USGS_cc4_lulc)[c(1:8)], collapse="+"), #families
                        sep = ""))
#that error message typically means you don’t have values for a variable represented in the model building process. 
water_gbm_lumped<- gbm(model,
             data=training, 
             distribution="bernoulli",
             n.trees=1111,
             shrinkage=0.1,
             interaction.depth=1,
             bag.fraction=0.50,
             #train.fraction=1,
             n.minobsinnode=10,
            cv.folds=10, # that error occurs when you have variable levels that exist in the test set which don't exist in the training set.
             keep.data=TRUE,
             class.stratify.cv=TRUE,
             verbose="CV")
#jpeg(paste("/Users/owner/Desktop/crypto/ec/","lumped_perf",'.jpg',sep=''))
gbm.perf(water_gbm_lumped)
save(water_gbm_lumped,file="/Users/owner/Desktop/crypto/ec/water_gbm_simple_Tavg.Rdata")
```


```{r}
best.iter<-gbm.perf(water_gbm_lumped)
gbm:::interact.gbm(water_gbm_lumped, data=USGS_cc4_lulc, i.var=c("Surface_runoff","Runoff"), best.iter)
gbm:::interact.gbm(water_gbm_lumped, data=USGS_cc4_lulc, i.var=c("Cultivated","Runoff"), best.iter)
```

#summary
```{r}
par(mar = c(5, 14, 1, 1))
# summary(
#  water_gbm_lumped , 
#   cBars = 10,
#   method = relative.influence, # also can use permutation.test.gbm
#   las = 2
#   )

summary(
 water_gbm_lumped , 
  cBars = 14,
  method = relative.influence, # also can use permutation.test.gbm
  las = 2
  )

#boxplot(USGS_cc4_lulc$lulc90~USGS_cc4_lulc$detect )
#boxplot(log(as.numeric(USGS_cc4_lulc$Surface_runoff )) ~USGS_cc4_lulc$Detect )

#percent positive in summer & winter
#length(USGS_cc4_lulc[USGS_cc4_lulc$detect==1&USGS_cc4_lulc$summer==1,1])/length(USGS_cc4_lulc[USGS_cc4_lulc$summer==1,1])
#length(USGS_cc4_lulc[USGS_cc4_lulc$detect==1&USGS_cc4_lulc$winter==1,1])/length(USGS_cc4_lulc[USGS_cc4_lulc$winter==1,1])
#print(water_gbm)

```

#n a ROC curve the true positive rate (Sensitivity) is plotted in function of the false positive rate (100-Specificity) for different cut-off points of a parameter.
```{r}
preds <- predict.gbm( water_gbm_lumped, test,n.trees=
                        #water_gbm$gbm.call$best.trees
                        class(gbm.perf(water_gbm_lumped)),
                      type="response")
calc.deviance(obs=as.numeric(test$Detect ) , pred=preds
              , calc.mean=TRUE
              )
roc_obj <- roc(test$Detect ,preds) #roc(response,predictor)
auc(roc_obj)
#jpeg(paste("/Users/owner/Desktop/crypto/ec/","roc_lumped",'.jpg',sep=''))
plot(roc_obj)
#dev.off()
plot(test$Detect ,preds)

predict_binary<-as.factor(ifelse(preds>0.5,1,0) )
test_obs<-as.factor(test$Detect)
confusionMatrix(predict_binary,test_obs)
```


#n a ROC curve the true positive rate (Sensitivity) is plotted in function of the false positive rate (100-Specificity) for different cut-off points of a parameter.
```{r}
preds <- predict.gbm( water_gbm_lumped, test,n.trees=
                        #water_gbm$gbm.call$best.trees
                        class(gbm.perf(water_gbm_lumped)),
                      type="response")
calc.deviance(obs=as.numeric(test$Detect ) , pred=preds
              , calc.mean=TRUE
              )
roc_obj <- roc(test$Detect ,preds) #roc(response,predictor)
auc(roc_obj)
#jpeg(paste("/Users/owner/Desktop/crypto/ec/","roc_lumped",'.jpg',sep=''))
plot(roc_obj)
#dev.off()
plot(test$Detect ,preds)

predict_binary<-as.factor(ifelse(preds>0.5,1,0) )
test_obs<-as.factor(test$Detect)
confusionMatrix(predict_binary,test_obs)
```

#Partial dependence plots. After the most relevant variables have been identified, the next step is to attempt to understand how the response variable changes based on these variables. 
#PDPs plot the change in the average predicted value as specified feature(s) vary over their marginal distribution. 
```{r}
jpeg(paste("/Users/owner/Desktop/crypto/ec/","relative_influence",'.jpg',sep=''))
par(mar = c(5, 11, 1, 1))
summary(
 water_gbm_lumped , 
  cBars = 11,
  method = relative.influence, # also can use permutation.test.gbm
  las = 2
  )
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/ec/","runoff",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "runoff", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
#+scale_y_continuous(labels = scales::dollar)
 #+scale_y_continuous(labels = scales::marginal effect )

dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/ec/","Tavg",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Tavg", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))

dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/ec/","Surface_runoff",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Surface_runoff", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()



jpeg(paste("/Users/owner/Desktop/crypto/ec/","Min_temperature",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Min_temperature", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/ec/","runoff",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "runoff", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()



jpeg(paste("/Users/owner/Desktop/crypto/ec/","Developed",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Developed", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/ec/","Forest",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Forest", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/ec/","Cultivated",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Cultivated", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))


dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/ec/","Forest_Runoff",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Forest_Runoff", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))

dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/ec/","Watershed_Area",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Watershed_Area", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
#partialPlot(boston.rf, pred.data = boston, x.var = "lstat",cex.lab=2.3)
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/ec/","Max_temperature",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Max_temperature", n.trees = water_gbm_lumped$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))


```

#PDP for interactive terms
```{r}
jpeg(paste("/Users/owner/Desktop/crypto/ec/","relative_influence_i",'.jpg',sep=''))
par(mar = c(5, 14, 1, 1))
summary(
 water_gbm_lumped , 
  cBars = 10,
  method = relative.influence, # also can use permutation.test.gbm
  las = 2
  )
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Surface_runoff_i",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Surface_runoff", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Runoff_Surface_runoff_i",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Runoff_Surface_runoff", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Temperature_Surface_runoff_i",'.jpg',sep=''))
water_gbm_lumped %>%
  partial(pred.var =  "Temperature_Surface_runoff", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Developed_Surface_runoff",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Developed_Surface_runoff", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Developed_Runoff",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Developed_Runoff", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Temperature_i",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Temperature", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Forest_Surface_runoff",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Forest_Surface_runoff", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Forest_Runoff",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Forest_Runoff", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Runoff_i",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Runoff", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Forest_i",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Forest", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Developed_i",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Developed", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Cultivated_i",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Cultivated", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
dev.off()

jpeg(paste("/Users/owner/Desktop/crypto/cry/","Watershed_Area_i",'.jpg',sep=''))
water_gbm_lumped_i %>%
  partial(pred.var =  "Watershed_Area", n.trees = water_gbm_lumped_i$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)+theme(axis.text=element_text(size=21),
        axis.title=element_text(size=24
                                #,face="bold"
                                ))
#partialPlot(boston.rf, pred.data = boston, x.var = "lstat",cex.lab=2.3)



```


```{r}
pretty.gbm.tree(water_gbm_lumped,i.tree=40)
```

```{r}
library(dismo)
data(Anguilla_train)
head(Anguilla_train)
angaus.tc5.lr01 <- gbm.step(data=Anguilla_train, gbm.x = 3:13, gbm.y = 2,
                        family = "bernoulli", tree.complexity = 5,                         learning.rate = 0.01, bag.fraction = 0.5)

angaus.tc5.lr005 <- gbm.step(data=Anguilla_train, gbm.x = 3:13, gbm.y = 2, 
                      family = "bernoulli", tree.complexity = 5,
                      learning.rate = 0.005, bag.fraction = 0.5)

find.int <- gbm.interactions(angaus.tc5.lr005)
find.int$interactions
```









#log
```{r}

USGS_cc4_lulc<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="TAVG"] <- "Temperature"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="storm"] <- "Surface_runoff"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc21"] <- "Developed_OpenSpace"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc22"] <- "Developed_LowIntensity"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc41"] <- "Forest"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc81"] <- "Pasture"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc82"] <- "Cultivated_Crops"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="detect"] <- "Detect"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="DrainageAreaMeasure.MeasureValue"] <- "Watershed_Area"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="runoff"] <- "log_runoff"

USGS_cc4_lulc_log<-USGS_cc4_lulc

USGS_cc4_lulc_log$log_runoff<-log(USGS_cc4_lulc_log$log_runoff)
#Partitioning the data and running a GBM

#Creating training and test sets
set.seed(123) #uses a fixed seed so that every call will result in the same output.

USGS_cc4_lulc_log<-USGS_cc4_lulc_log[,c(36:38,29,26,3:7)]

intrain<-createDataPartition(y=USGS_cc4_lulc_log$Detect ,
                             p=0.8,
                             list=FALSE)
#class(intrain) "matrix"

test<-USGS_cc4_lulc_log[-intrain,]
training<-USGS_cc4_lulc_log[intrain,]

dim(test)
dim(training)
```


<!-- #add summer and winter -->
<!-- ```{r} -->

<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$summer<-0 -->
<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws[USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$month==6|USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$month==7|USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$month ==8 ,]$summer<-1 -->
<!-- summary(USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$summer) -->

<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$winter<-0 -->
<!-- USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws[USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$month==12|USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$month==1|USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$month ==2 ,]$winter<-1 -->
<!-- summary(USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws$winter) -->

<!-- ``` -->

#run the model
```{r}

USGS_cc4_lulc<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws

#Partitioning the data and running a GBM

#Creating training and test sets
set.seed(123) #uses a fixed seed so that every call will result in the same output.

USGS_cc4_lulc<-USGS_cc4_lulc[,c(36:38,29,26,2:16)]

intrain<-createDataPartition(y=USGS_cc4_lulc$detect ,
                             p=0.8,
                             list=FALSE)
#class(intrain) "matrix"

test<-USGS_cc4_lulc[-intrain,]
training<-USGS_cc4_lulc[intrain,]

dim(test)
dim(training)
```


<!-- #run the model -->
<!-- ```{r} -->

<!-- #Partitioning the data and running a GBM -->

<!-- #Creating training and test sets -->
<!-- set.seed(21) #uses a fixed seed so that every call will result in the same output. -->
<!-- intrain<-createDataPartition(y=USGS_cc4_lu21$detect , -->
<!--                              p=0.8, -->
<!--                              list=FALSE) -->
<!-- #class(intrain) "matrix" -->

<!-- test<-USGS_cc4_lu21[-intrain,] -->
<!-- training<-USGS_cc4_lu21[intrain,] -->

<!-- ``` -->

<!-- #fake big model -->
<!-- ```{r} -->
<!-- load("/Users/owner/Desktop/crypto/USGS_cc4_lu21.Rdata") -->
<!-- #USGS_cc4_lu21[,24]<-as.numeric(USGS_cc4_lu21[,24]) -->
<!-- fake<-USGS_cc4_lu21[,c(11,14,21,24)] -->
<!-- fake$DrainageAreaMeasure.MeasureValue2<-(runif(125, min= min(fake$DrainageAreaMeasure.MeasureValue) , max=max(fake$DrainageAreaMeasure.MeasureValue)) ) -->
<!-- fake$runoff2 <-(runif(125, min= min(fake$runoff ) , max=max(fake$runoff )) ) -->
<!-- fake$percent2<-(runif(125, min= 0 , max= 1) ) -->
<!-- head(fake) -->
<!-- set.seed(21) #uses a fixed seed so that every call will result in the same output. -->
<!-- intrain<-createDataPartition(y=fake$detect , -->
<!--                              p=0.8, -->
<!--                              list=FALSE) -->
<!-- #class(intrain) "matrix" -->

<!-- test<-fake[-intrain,] -->
<!-- training<-fake[intrain,] -->
<!-- model<-as.formula(paste(colnames(fake)[1], "~",      paste(colnames(fake)[c(2:7)], collapse="+"), #families -->
<!--                         sep = "")) -->
<!-- #that error message typically means you don’t have values for a variable represented in the model building process.  -->
<!-- water_gbm<- gbm(model, -->
<!--              data=training,  -->
<!--              distribution="bernoulli", -->
<!--              n.trees=100, -->
<!--              shrinkage=0.1, -->
<!--              interaction.depth=1, -->
<!--              bag.fraction=0.50, -->
<!--              #train.fraction=1, -->
<!--              n.minobsinnode=10, -->
<!--              cv.folds=10, -->
<!--              keep.data=TRUE, -->
<!--              class.stratify.cv=TRUE, -->
<!--              verbose="CV") -->

<!-- ``` -->

<!-- #the model -->
<!-- ```{r} -->
<!-- # generalized boosted regression model formula -->
<!-- #model<-as.formula(fake_glm_data$cry_detect~fake_glm_data$runoff+fake_glm_data$summer+fake_glm_data$watershed_size+fake_glm_data$urb+fake_glm_data$temp ) -->
<!-- USGS_cc4_lu21[,24]<-as.numeric(USGS_cc4_lu21[,24]) -->
<!-- model<-as.formula(paste(colnames(USGS_cc4_lu21)[11], "~",      paste(colnames(USGS_cc4_lu21)[c(14,21,24)], collapse="+"), #families -->
<!--                         sep = "")) -->
<!-- #that error message typically means you don’t have values for a variable represented in the model building process.  -->
<!-- water_gbm<- gbm(model, -->
<!--              data=training,  -->
<!--              distribution="bernoulli", -->
<!--              n.trees=100, -->
<!--              shrinkage=0.1, -->
<!--              interaction.depth=1, -->
<!--              bag.fraction=0.50, -->
<!--              #train.fraction=1, -->
<!--              n.minobsinnode=10, -->
<!--              cv.folds=10, -->
<!--              keep.data=TRUE, -->
<!--              class.stratify.cv=TRUE, -->
<!--              verbose="CV") -->

<!-- ``` -->

#look at the data
```{r}
boxplot(as.numeric(USGS_cc4_lulc$detect)~(USGS_cc4_lulc$GAGE_ID))
boxplot(log(as.numeric(USGS_cc4_lulc$runoff ))~(USGS_cc4_lulc$detect ))
boxplot((as.numeric(USGS_cc4_lulc$lulc21 ))~(USGS_cc4_lulc$detect ))
boxplot((as.numeric(USGS_cc4_lulc$lulc41 ))~(USGS_cc4_lulc$detect ))
boxplot((as.numeric(USGS_cc4_lulc$lulc81 ))~(USGS_cc4_lulc$detect ))
boxplot(log(as.numeric(USGS_cc4_lulc$DrainageAreaMeasure.MeasureValue ))~(USGS_cc4_lulc$detect ))
```

<!-- #as.factor -->
<!-- ```{r} -->
<!-- for (c in c(5,9,13,17,21,25,29,33,46,53)) -->
<!-- { -->
<!--  print (test[,c]%in%training[,c]) -->
<!--   print (class(test[,c])) -->
<!--   training[,c]<-as.factor(training[,c]) -->
<!--   test[,c]<-as.factor(test[,c]) -->
<!-- } -->
<!-- #test -->
<!-- #training -->

<!-- #43], "~",      paste(colnames(USGS_cc4_lulc)[c(5,9,13,17,21,25,29,33,46,53 -->
<!-- ``` -->


#the model
```{r}
# generalized boosted regression model formula
#model<-as.formula(fake_glm_data$cry_detect~fake_glm_data$runoff+fake_glm_data$summer+fake_glm_data$watershed_size+fake_glm_data$urb+fake_glm_data$temp )
#USGS_cc4_lu21[,24]<-as.numeric(USGS_cc4_lu21[,24])
# summary(lm(as.numeric(USGS_cc4_lulc$detect))~
#              USGS_cc4_lulc$lulc41 + 
#              USGS_cc4_lulc$lulc81 + 
#              USGS_cc4_lulc$lulc90 + 
#              USGS_cc4_lulc$lulc21 + 
#              USGS_cc4_lulc$lulc11 + 
#              USGS_cc4_lulc$lulc43 +
#              USGS_cc4_lulc$lulc52 +
#              USGS_cc4_lulc$lulc22 +
#              USGS_cc4_lulc$lulc23 +
#              USGS_cc4_lulc$DrainageAreaMeasure.MeasureValue + 
#              USGS_cc4_lulc$runoff+USGS_cc4_lulc$storm
#             # USGS_cc4_lulc$summer+USGS_cc4_lulc$winter
#            )
#USGS_cc4_lulc[,9]<-as.factor(USGS_cc4_lulc[,9])

model<-as.formula(paste(colnames(USGS_cc4_lulc)[5], "~",      paste(colnames(USGS_cc4_lulc)[c(1:4,6:20)], collapse="+"), #families
                        sep = ""))
#that error message typically means you don’t have values for a variable represented in the model building process. 
water_gbm<- gbm(model,
             data=training, 
             distribution="bernoulli",
             n.trees=400,
             shrinkage=0.1,
             interaction.depth=1,
             bag.fraction=0.50,
             #train.fraction=1,
             n.minobsinnode=10,
             cv.folds=10, # that error occurs when you have variable levels that exist in the test set which don't exist in the training set.
             keep.data=TRUE,
             class.stratify.cv=TRUE,
             verbose="CV")
gbm.perf(water_gbm)

```


<!-- #the model (clean input data) -->
<!-- ```{r} -->
<!-- # generalized boosted regression model formula -->
<!-- #model<-as.formula(fake_glm_data$cry_detect~fake_glm_data$runoff+fake_glm_data$summer+fake_glm_data$watershed_size+fake_glm_data$urb+fake_glm_data$temp ) -->
<!-- #USGS_cc4_lu21[,24]<-as.numeric(USGS_cc4_lu21[,24]) -->
<!-- #summary(lm(as.numeric(USGS_cc4_lulc$detect)~USGS_cc4_lulc$lulc41 + USGS_cc4_lulc$lulc81 + USGS_cc4_lulc$lulc90 + USGS_cc4_lulc$lulc21 + USGS_cc4_lulc$lulc11 + USGS_cc4_lulc$lulc43 +USGS_cc4_lulc$DrainageAreaMeasure.MeasureValue + USGS_cc4_lulc$runoff)) -->
<!-- #USGS_cc4_lulc<-USGS_cc4_lulc[,c(5,9,13,17,21,25,29,33,43,46,53)] -->
<!-- model<-as.formula(paste(colnames(USGS_cc4_lulc)[9], "~",      paste(colnames(USGS_cc4_lulc)[c(1:8,10,11 -->
<!--                                                                                                )], collapse="+"), #families -->
<!--                         sep = "")) -->
<!-- #that error message typically means you don’t have values for a variable represented in the model building process. -->
<!-- water_gbm<- gbm(model, -->
<!--              data=training, -->
<!--              distribution="bernoulli", -->
<!--              n.trees=100, -->
<!--              shrinkage=0.1, -->
<!--              interaction.depth=1, -->
<!--              bag.fraction=0.50, -->
<!--              #train.fraction=1, -->
<!--              n.minobsinnode=10, -->
<!--              cv.folds=10, # that error occurs when you have variable levels that exist in the test set which don't exist in the training set. -->
<!--              keep.data=TRUE, -->
<!--              class.stratify.cv=TRUE, -->
<!--              verbose="CV") -->

<!-- ``` -->


<!-- ```{r} -->
<!-- for (c in c(5,9,13,17,21,25,29,33,43,46,53)) -->
<!-- { -->
<!--  print (test[,c]%in%training[,c]) -->
<!--   print (class(test[,c])) -->
<!-- } -->
<!-- #test -->
<!-- #training -->

<!-- #43], "~",      paste(colnames(USGS_cc4_lulc)[c(5,9,13,17,21,25,29,33,46,53 -->
<!-- ``` -->


#lulc41 forest, #lulc90 wetland #21 developed #52	Shrub #11	Open Water #81	Pasture
```{r}
summary(water_gbm)
#boxplot(USGS_cc4_lulc$lulc90~USGS_cc4_lulc$detect )
#boxplot(log(as.numeric(USGS_cc4_lulc$runoff)) ~USGS_cc4_lulc$detect )

#percent positive in summer & winter
length(USGS_cc4_lulc[USGS_cc4_lulc$detect==1&USGS_cc4_lulc$summer==1,1])/length(USGS_cc4_lulc[USGS_cc4_lulc$summer==1,1])
length(USGS_cc4_lulc[USGS_cc4_lulc$detect==1&USGS_cc4_lulc$winter==1,1])/length(USGS_cc4_lulc[USGS_cc4_lulc$winter==1,1])


```

#predict on training
```{r}
gbmTrainPredictions = predict(object =water_gbm  ,
                              #newdata = train,
                             # n.trees = 100,
                              type = "response")
calibrate.plot(training$detect , gbmTrainPredictions, xlim=c(0,1), ylim=c(0,1))
```

#predict on test
```{r}
gbmTrainPredictions = predict(object =water_gbm  ,
                              newdata = test,
                             # n.trees = 100,
                              type = "response")

plot(test$detect,gbmTrainPredictions,ylim=c(0,1))
```


#run the model (only run mean LULC>10% (lulc21,22,41,81,82))
```{r}

USGS_cc4_lulc<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="TAVG"] <- "Temperature"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="storm"] <- "Surface_runoff"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc21"] <- "Developed_OpenSpace"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc22"] <- "Developed_LowIntensity"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc41"] <- "Forest"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc81"] <- "Pasture"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc82"] <- "Cultivated_Crops"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="detect"] <- "Detect"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="DrainageAreaMeasure.MeasureValue"] <- "Watershed_Area"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="runoff"] <- "Runoff"

#Partitioning the data and running a GBM

#Creating training and test sets
set.seed(123) #uses a fixed seed so that every call will result in the same output.

USGS_cc4_lulc<-USGS_cc4_lulc[,c(36:38,29,26,3:7)]

intrain<-createDataPartition(y=USGS_cc4_lulc$Detect ,
                             p=0.8,
                             list=FALSE)
#class(intrain) "matrix"

test<-USGS_cc4_lulc[-intrain,]
training<-USGS_cc4_lulc[intrain,]

dim(test)
dim(training)
```
#plot hist
```{r}
for (v in c(1:length(colnames(USGS_cc4_lulc))))
            {
  if (!class(USGS_cc4_lulc[,v])=="character")
  {
    jpeg(paste("/Users/owner/Desktop/crypto/cry/",colnames(USGS_cc4_lulc)[v],'.jpg',sep=''))

  hist(USGS_cc4_lulc[,v],xlab=colnames(USGS_cc4_lulc)[v],main="",ylim=c(0,100))
dev.off()
  }     
      }
```


#the model
```{r}
# generalized boosted regression model formula
#model<-as.formula(fake_glm_data$cry_detect~fake_glm_data$runoff+fake_glm_data$summer+fake_glm_data$watershed_size+fake_glm_data$urb+fake_glm_data$temp )
#USGS_cc4_lu21[,24]<-as.numeric(USGS_cc4_lu21[,24])

#USGS_cc4_lulc[,9]<-as.factor(USGS_cc4_lulc[,9])

model<-as.formula(paste(colnames(USGS_cc4_lulc)[5], "~",      paste(colnames(USGS_cc4_lulc)[c(1:4,6:10)], collapse="+"), #families
                        sep = ""))
#that error message typically means you don’t have values for a variable represented in the model building process. 
water_gbm<- gbm(model,
             data=training, 
             distribution="bernoulli",
             n.trees=600,
             shrinkage=0.1,
             interaction.depth=1,
             bag.fraction=0.50,
             #train.fraction=1,
             n.minobsinnode=10,
             cv.folds=10, # that error occurs when you have variable levels that exist in the test set which don't exist in the training set.
             keep.data=TRUE,
             class.stratify.cv=TRUE,
             verbose="CV")
# find index for n trees with minimum CV error
min_MSE <- which.min(water_gbm$cv.error)
print(min_MSE)
gbm.perf(water_gbm
         #,oobag.curve =TRUE
         ) #the optimal number of boosting iterations 
#an independent test set (test), out-of-bag estima- tion (OOB), and v-fold cross validation (cv)
#overlay a right y-axis is added to the training and test error plot and the estimated cumulative improvement in the loss function is plotted versus the iteration number.
#gbm.perf(water_gbm,method = "OOB")

#print the first and last trees
#plot(water_gbm$trees,water_gbm$train.error)
```

```{r}
summary(lm(USGS_cc4_lulc$Detect~USGS_cc4_lulc$Runoff+USGS_cc4_lulc$Surface_runoff+USGS_cc4_lulc$Temperature+USGS_cc4_lulc$Watershed_Area+USGS_cc4_lulc$Developed_OpenSpace+USGS_cc4_lulc$Developed_LowIntensity+USGS_cc4_lulc$Forest+USGS_cc4_lulc$Cultivated_Crops+USGS_cc4_lulc$Pasture))
```


#lulc41 forest, #lulc90 wetland #21 developed #52	Shrub #11	Open Water #81	Pasture
```{r}
par(mar = c(5, 11, 1, 1))
summary(
 water_gbm, 
  cBars = 10,
  method = relative.influence, # also can use permutation.test.gbm
  las = 2
  )
#boxplot(USGS_cc4_lulc$lulc90~USGS_cc4_lulc$detect )
boxplot(log(as.numeric(USGS_cc4_lulc$Surface_runoff )) ~USGS_cc4_lulc$Detect )

#percent positive in summer & winter
length(USGS_cc4_lulc[USGS_cc4_lulc$detect==1&USGS_cc4_lulc$summer==1,1])/length(USGS_cc4_lulc[USGS_cc4_lulc$summer==1,1])
length(USGS_cc4_lulc[USGS_cc4_lulc$detect==1&USGS_cc4_lulc$winter==1,1])/length(USGS_cc4_lulc[USGS_cc4_lulc$winter==1,1])
print(water_gbm)

```

#n a ROC curve the true positive rate (Sensitivity) is plotted in function of the false positive rate (100-Specificity) for different cut-off points of a parameter.
```{r}
preds <- predict.gbm( water_gbm, test,n.trees=
                        #water_gbm$gbm.call$best.trees
                        class(gbm.perf(water_gbm)),
                      type="response")
calc.deviance(obs=as.numeric(test$Detect ) , pred=preds
              , calc.mean=TRUE
              )
roc_obj <- roc(test$Detect ,preds)
auc(roc_obj)
plot(roc_obj)
plot(test$Detect ,preds)
predict_binary<-as.factor(ifelse(preds>0.5,1,0) )
test_obs<-as.factor(test$Detect)
confusionMatrix(predict_binary,test_obs)
```


#Partial dependence plots. After the most relevant variables have been identified, the next step is to attempt to understand how the response variable changes based on these variables. 
#PDPs plot the change in the average predicted value as specified feature(s) vary over their marginal distribution. 
```{r}
summary(water_gbm)
 water_gbm %>%
  partial(pred.var =  "Surface_runoff", n.trees = water_gbm$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)
#+scale_y_continuous(labels = scales::dollar)
 #+scale_y_continuous(labels = scales::marginal effect )

water_gbm %>%
  partial(pred.var =  "Forest", n.trees = water_gbm$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)
water_gbm %>%
  partial(pred.var =  "Cultivated_Crops", n.trees = water_gbm$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)
 water_gbm %>%
   partial(pred.var =  "Runoff", n.trees = water_gbm$n.trees, grid.resolution = 100) %>%
   autoplot(rug = TRUE, train = training)
water_gbm %>%
  partial(pred.var =  "Watershed_Area", n.trees = water_gbm$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)
water_gbm %>%
  partial(pred.var =  "Developed_OpenSpace", n.trees = water_gbm$n.trees, grid.resolution = 100) %>%autoplot(rug = TRUE, train = training)
water_gbm %>%
  partial(pred.var =  "Temperature", n.trees = water_gbm$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)
water_gbm %>%
  partial(pred.var =  "Pasture", n.trees = water_gbm$n.trees, grid.resolution = 100) %>%
  autoplot(rug = TRUE, train = training)
```











#log
```{r}

USGS_cc4_lulc<-USGS_cc4_lulc_CntlPlains_EastHghlnds_NorthEast_cry_ws
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="TAVG"] <- "Temperature"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="storm"] <- "Surface_runoff"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc21"] <- "Developed_OpenSpace"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc22"] <- "Developed_LowIntensity"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc41"] <- "Forest"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc81"] <- "Pasture"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="lulc82"] <- "Cultivated_Crops"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="detect"] <- "Detect"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="DrainageAreaMeasure.MeasureValue"] <- "Watershed_Area"
colnames(USGS_cc4_lulc)[colnames( USGS_cc4_lulc)=="runoff"] <- "log_runoff"

USGS_cc4_lulc_log<-USGS_cc4_lulc

USGS_cc4_lulc_log$log_runoff<-log(USGS_cc4_lulc_log$log_runoff)
#Partitioning the data and running a GBM

#Creating training and test sets
set.seed(123) #uses a fixed seed so that every call will result in the same output.

USGS_cc4_lulc_log<-USGS_cc4_lulc_log[,c(36:38,29,26,3:7)]

intrain<-createDataPartition(y=USGS_cc4_lulc_log$Detect ,
                             p=0.8,
                             list=FALSE)
#class(intrain) "matrix"

test<-USGS_cc4_lulc_log[-intrain,]
training<-USGS_cc4_lulc_log[intrain,]

dim(test)
dim(training)
```


#the model
```{r}
model_log<-as.formula(paste(colnames(USGS_cc4_lulc_log)[5], "~",      paste(colnames(USGS_cc4_lulc_log)[c(1:4,6:10)], collapse="+"), #families
                        sep = ""))
#that error message typically means you don’t have values for a variable represented in the model building process. 
water_gbm_log<- gbm(model_log,
             data=training, 
             distribution="bernoulli",
             n.trees=600,
             shrinkage=0.1,
             interaction.depth=1,
             bag.fraction=0.50,
             #train.fraction=1,
             n.minobsinnode=10,
             cv.folds=10, # that error occurs when you have variable levels that exist in the test set which don't exist in the training set.
             keep.data=TRUE,
             class.stratify.cv=TRUE,
             verbose="CV")

gbm.perf(water_gbm_log
         #,oobag.curve =TRUE
         ) #the optimal number of boosting iterations 
#an independent test set (test), out-of-bag estima- tion (OOB), and v-fold cross validation (cv)
#overlay a right y-axis is added to the training and test error plot and the estimated cumulative improvement in the loss function is plotted versus the iteration number.
#gbm.perf(water_gbm,method = "OOB")

#print the first and last trees
#plot(water_gbm$trees,water_gbm$train.error)
```


#lulc41 forest, #lulc90 wetland #21 developed #52	Shrub #11	Open Water #81	Pasture
```{r}
par(mar = c(5, 11, 1, 1))
summary(
 water_gbm_log, 
  cBars = 10,
  method = relative.influence, # also can use permutation.test.gbm
  las = 2
  )

print(water_gbm_log)

```




#find high runoff sampled watershed
```{r}
mean_runoff<-aggregate(USGS_cc4[,"runoff"], list(USGS_cc4$MonitoringLocationIdentifier ), mean)

median_runoff<-aggregate(USGS_cc4[,"runoff"], list(USGS_cc4$MonitoringLocationIdentifier ), median )

n_obs<-as.data.frame(table(USGS_cc4$ MonitoringLocationIdentifier))

mean_runoff_obs<-merge(mean_runoff,n_obs,by.x= "Group.1",by.y= "Var1")

mean_runoff_obs[mean_runoff_obs$x>3,]
mean_runoff_obs[mean_runoff_obs$x<1,]

```


#find high concentration sampled watershed (non detect sample w conc?)
```{r}
USGS_cc4$ResultMeasureValue<-as.numeric(USGS_cc4$ResultMeasureValue)
mean_conc<-aggregate(USGS_cc4[,"ResultMeasureValue" ], list(USGS_cc4$MonitoringLocationIdentifier), mean)

#median_runoff<-aggregate(USGS_cc4[,"runoff"], list(USGS_cc4$MonitoringLocationIdentifier ), median )

n_obs<-as.data.frame(table(USGS_cc4$ MonitoringLocationIdentifier))

mean_conc_obs<-merge(mean_conc,n_obs,by.x= "Group.1",by.y= "Var1")

mean_conc_obs[mean_conc_obs$x>60,]
mean_conc_obs[mean_conc_obs$x<30,]

USGS_04087214<-USGS_cc4[USGS_cc4$MonitoringLocationIdentifier=="USGS-04087214",]
par(mar=c(3,3,3,3))
plot(USGS_04087214$runoff ,USGS_04087214$ResultMeasureValue )
```

```{r}
par(mar=c(1,1,1,1))
plot(USGS_04087214$runoff ,USGS_04087214$ResultMeasureValue )
par(mar=c(1.5,1.5,1.5,1.5))
plot(USGS_04087214$month ,USGS_04087214$ResultMeasureValue )
```


#need to get Q in previous 48 days 
```{r}
USGS_cc4$storm<-" "
for (i in c(1:length(USGS_cc4$USGS)))
#for (i in c(41:59))
{
siteNo <- as.character(USGS_cc4[i,]$USGS)
pCode <- "00060" #discharge
start.date <- as.character(as.Date(USGS_cc4[i,]$ActivityStartDate,format=("%Y-%m-%d"))-48)
end.date <- as.character(USGS_cc4[i,]$ActivityStartDate)

if (nrow(readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = start.date,
                     endDate = end.date))>0)
{
  print(nrow(readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = start.date,
                     endDate = end.date)))
  USGS_cc4[i,]$Q<- readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = end.date,
                     endDate = end.date)$X_00060_00003
  
 st<- paste("Q_48days",i,sep='_')
  st<- readNWISdv(siteNumbers = siteNo,
                     parameterCd = pCode,
                     startDate = start.date,
                     endDate = end.date)
  print (paste("Q_48days",i,sep='_'))
  #hysep(s1372500$X105182_00060_00003,Dates=2/1/2019,da=1)
  #head(s1372500)
  #agency_cd site_no datetime X105182_00060_00003 X105182_00060_00003_cd
#1      USGS 1372500 2/1/2019                 399                      P
test<-hysep(as.integer(st$X_00060_00003),Dates=
              #2008-09-27
              #9/27/2008
              st$Date
            ,da=1)
print(test)
USGS_cc4[i,]$storm<-(1-test$BaseQ[49]/test$Flow[49])
}
else
{
 USGS_cc4[i,]$storm<-NA 
 USGS_cc4[i,]$Q<-NA 
}
  }
head(USGS_cc4)
plot(USGS_cc4$storm,as.numeric(USGS_cc4$runoff))
                                
```


#fix unit
```{r}
#summary(USGS_cc2$ResultMeasure.MeasureUnitCode)
               #/100L        #/l #/l                 % ocyst/100L oocyst/10L 
#       100          0          0          0          3         63        272  

#oocyst/10L
USGS_cc4$oocyst10L<-" "

USGS_cc4<-USGS_cc4[USGS_cc4$ResultMeasure.MeasureUnitCode=="oocyst/10L"|USGS_cc4$ResultMeasure.MeasureUnitCode=="ocyst/100L",]


summary(USGS_cc4$ResultMeasureValue)
i<-which(USGS_cc4$ResultMeasure.MeasureUnitCode=="oocyst/10L")
{
  USGS_cc4[i,]$oocyst10L<-as.numeric(as.character(USGS_cc4[i,]$ResultMeasureValue))
}

summary(as.numeric(USGS_cc4$oocyst10L))

j<-which(USGS_cc4$ResultMeasure.MeasureUnitCode=="ocyst/100L")
{
  USGS_cc4[j,]$oocyst10L<-as.numeric(as.character(USGS_cc4[j,]$ResultMeasureValue))/10
}
summary(as.numeric(USGS_cc4$oocyst10L))
```

```{r}
plot(USGS_cc4$runoff,USGS_cc4$oocyst10L)
plot(USGS_cc4$nlcd_2,USGS_cc4$oocyst10L)
plot(USGS_cc4$nlcd_8,USGS_cc4$oocyst10L)
plot(USGS_cc4$storm ,USGS_cc4$oocyst10L)
#plot(USGS_cc4$month ,USGS_cc4$oocyst10L)
boxplot(as.numeric(USGS_cc4$oocyst10L)~USGS_cc4$month,ylab="oocyst/10L",xlab="month")
boxplot(as.numeric(USGS_cc4$oocyst10L)~USGS_cc4$nlcd_2,ylab="oocyst/10L",xlab="month")
#plot(USGS_cc4$month,USGS_cc4$detect)
#plot(USGS_cc4[USGS_cc4$month=="6"|USGS_cc4$month=="7"|USGS_cc4$month=="8",]$runoff ,USGS_cc4[USGS_cc4$month=="6"|USGS_cc4$month=="7"|USGS_cc4$month=="8",]$oocyst10L)
plot(USGS_cc4$MonitoringLocationIdentifier,USGS_cc4$oocyst10L,ylab="oocyst/10L")
boxplot(as.numeric(USGS_cc4$oocyst10L)~USGS_cc4$MonitoringLocationIdentifier,ylab="oocyst/10L")
plot((USGS_cc4$DrainageAreaMeasure.MeasureValue),USGS_cc4$oocyst10L,log="x")
boxplot(as.numeric(USGS_cc4$oocyst10L)~USGS_cc4$DrainageAreaMeasure.MeasureValue ,ylab="oocyst/10L")
```

```{r}
USGS_cc4$oocyst10L<-as.numeric(USGS_cc4$oocyst10L)
USGS_cc4$runoff <-as.numeric(USGS_cc4$runoff )
USGS_cc4$DrainageAreaMeasure.MeasureValue<-as.numeric(USGS_cc4$DrainageAreaMeasure.MeasureValue)
USGS_cc4$storm<-as.numeric(USGS_cc4$storm)
USGS_cc4$nlcd_2_8<-as.numeric(USGS_cc4$nlcd_2+USGS_cc4$nlcd_8)
summary(lm(USGS_cc4$oocyst10L~USGS_cc4$runoff+USGS_cc4$DrainageAreaMeasure.MeasureValue+USGS_cc4$storm+USGS_cc4$nlcd_8))
```

```{r}
small<-USGS_cc4[USGS_cc4$DrainageAreaMeasure.MeasureValue<=25,]
plot(small$runoff,small$oocyst10L)
boxplot(as.numeric(small$oocyst10L)~small$MonitoringLocationIdentifier)
```


```{r}
write.csv(USGS_cc4,file="USGS_cc4.csv")
```


```{r}
USGS_cc2<-USGS_cc2[is.na(USGS_cc2$Q)==F,]

as.data.frame(table(USGS_cc2$MonitoringLocationIdentifier))
summary(USGS_cc2$ResultMeasureValue)
plot(USGS_cc2$Q,USGS_cc2$ResultMeasureValue)

USGS_01553500<-USGS_cc2[USGS_cc2$MonitoringLocationIdentifier=="USGS-01553500",]
plot(USGS_01553500$Q,USGS_01553500$ResultMeasureValue)

#count[count$Freq>09,]

USGS_03036500<-USGS_cc2[USGS_cc2$MonitoringLocationIdentifier=="USGS-03036500",]
plot(USGS_03036500$Q,USGS_03036500$ResultMeasureValue)

```

